---
title: "Gumbel Simulations"
author: '25062476'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ismev) # Coles Package
library(evmissing) # Simpson & Northrop package 
library(readr) # for csv and things 
library(ggplot2)
library(xtable)

library(dplyr)
library(knitr)
library(kableExtra)
```

## This is generally all the code used for Chapter 3 (Simulations)
Includes several cases that I have tested for the Gumbel domain of attraction using independent samples with monthly means and standard deviations. 
Idea for this Markdown is to isolate the code so that we will be using some distribution (The Gaussian) that lies in the domain of attraction of Gumbel. 

## Functions required to run the other chunks of this R markdown document 
```{r Functions for running the rest of this R markdown document}
x_samples <- function(n_samples, mu, sigma) {
  month_days  <- c(31,28,31,30,31,30,31,31,30,31,30,31)
  month_index <- rep(1:12, times = month_days)
  month_index <- rep(month_index, n_samples/365)
  rnorm(length(month_index), mean = mu[month_index], sd = sigma[month_index])
}

block_max <- function(x, block_size) {
  as.numeric(tapply(x, (seq_along(x) - 1) %/% block_size, max))
}

delta_rl_CI <- function(T, fit) {
  mu    <- fit$mle[1]
  sigma <- fit$mle[2]
  xi    <- fit$mle[3]
  V     <- fit$cov
  y     <- -log(1 - 1/T)

  if (abs(xi) < 1e-6) {
    rl     <- mu - sigma * log(y)
    var_rl <- V[1,1] + (log(y))^2 * V[2,2] - 2 * log(y) * V[1,2]
  } else {
    rl      <- mu + (sigma/xi) * (y^(-xi) - 1)
    grad_rl <- c(1, (y^(-xi) - 1)/xi, sigma * (1 - y^(-xi) - xi * y^(-xi) * log(y))/xi^2)
    var_rl  <- as.numeric(t(grad_rl) %*% V %*% grad_rl)
  }

  se_rl <- sqrt(var_rl)
  data.frame(ReturnLevel = rl, Lower = rl - 1.96 * se_rl, Upper = rl + 1.96 * se_rl)
}

delta_prof <- function(GEV_samples, n_samples, q_samples, T, mu, sigma) {

  # True return level from large sample (OR REFERENCE RETURN LEVELS) more correct! 
  x_true  <- x_samples(q_samples, mu, sigma)
  bm_true <- block_max(x_true, 365)
  true_rl <- as.numeric(quantile(bm_true, probs = 1 - 1/T))

  cat("True return level (T =", T, "):", round(true_rl, 4), "\n")

  results <- data.frame(
    delta_rl    = numeric(n_samples),
    delta_lower = numeric(n_samples),
    delta_upper = numeric(n_samples),
    prof_rl     = numeric(n_samples),
    prof_lower  = numeric(n_samples),
    prof_upper  = numeric(n_samples)
  )

  delta_fail     <- 0L
  profile_fail   <- 0L
  profileci_fail <- 0L

  for (i in seq_len(n_samples)) {

    if (i %% max(1, n_samples %/% 4) == 0)
      cat(round(100 * i / n_samples), "% done\n")

    x_dummy  <- x_samples(GEV_samples, mu, sigma)
    bm_dummy <- block_max(x_dummy, 365)

    # Delta method
    fit_delta <- tryCatch(
      suppressWarnings(gev.fit(bm_dummy, show = FALSE)),
      error = function(e) NULL
    )

    if (is.null(fit_delta)) {
      delta_fail <- delta_fail + 1L
      results[i, c("delta_rl", "delta_lower", "delta_upper")] <- NA
    } else {
      ci <- delta_rl_CI(T, fit_delta)
      results[i, "delta_rl"]    <- ci$ReturnLevel
      results[i, "delta_lower"] <- ci$Lower
      results[i, "delta_upper"] <- ci$Upper
    }

    # Profile likelihood
    fit_prof <- tryCatch(
      suppressWarnings(gev_mle(bm_dummy, 1)),
      error = function(e) NULL
    )

    if (is.null(fit_prof) || any(is.na(fit_prof))) {
      profile_fail <- profile_fail + 1L
      results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
    } else {
      prof_return <- gev_return(fit_prof, m = T)
      prof_conf   <- tryCatch(
        suppressWarnings(confint(prof_return, profile = TRUE)),
        error = function(e) NULL
      )

      if (is.null(prof_conf)) {
        profileci_fail <- profileci_fail + 1L
        results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
      } else {
        results[i, "prof_rl"]    <- as.numeric(prof_return)
        results[i, "prof_lower"] <- prof_conf[1, 1]
        results[i, "prof_upper"] <- prof_conf[1, 2]
      }
    }
  }

  # Summary dataframe to collect all the info I needed 
  summary_df <- data.frame(
    ReturnPeriod = T,
    true_rl      = true_rl,

    delta_rl   = mean(results$delta_rl,  na.rm = TRUE),
    prof_rl    = mean(results$prof_rl,   na.rm = TRUE),

    RMSE_delta = sqrt(mean((results$delta_rl - true_rl)^2, na.rm = TRUE)),
    RMSE_prof  = sqrt(mean((results$prof_rl  - true_rl)^2, na.rm = TRUE)),

    Bias_delta     = mean(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof      = mean(results$prof_rl  - true_rl, na.rm = TRUE),

    Bias_delta_med = median(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof_med  = median(results$prof_rl  - true_rl, na.rm = TRUE),

    Coverage_delta = mean(results$delta_lower <= true_rl &
                          results$delta_upper >= true_rl, na.rm = TRUE),
    Coverage_prof  = mean(results$prof_lower  <= true_rl &
                          results$prof_upper  >= true_rl, na.rm = TRUE),
    
    delta_width    = mean(results$delta_upper - results$delta_lower, na.rm=TRUE),
    prof_width     = mean(results$prof_upper - results$prof_lower, na.rm=TRUE)
    
  )

  list(
    summary            = summary_df,
    raw                = results,       
    true_rl            = true_rl,
    delta_failures     = delta_fail,
    profile_failures   = profile_fail,
    profileci_failures = profileci_fail,
    delta_failure_rate   = delta_fail     / n_samples,
    profile_failure_rate = profile_fail   / n_samples
  )
}

coverage_plots <- function(raw, true_rl) {
  
  raw     <- raw
  true_rl <- true_rl
  
  n <- nrow(raw)
  raw$iteration <- seq_len(n)
  
  raw$delta_covers <- raw$delta_lower <= true_rl & raw$delta_upper >= true_rl
  raw$prof_covers  <- raw$prof_lower  <= true_rl & raw$prof_upper  >= true_rl
  
  # ── Delta plot ────────────────────────────────────────────────────────────────
  p_delta <- ggplot(raw, aes(x = iteration, colour = delta_covers)) +
  
    geom_linerange(aes(ymin = delta_lower, ymax = delta_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = delta_rl), shape = 16, size = 1.5, alpha = 0.5) +
  
    scale_colour_manual(
      name   = "Covers True Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", hjust = 0.5),
      legend.position  = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin    = margin(4, 6, 4, 6)
    )
  
  # ── Profile plot ──────────────────────────────────────────────────────────────
  p_prof <- ggplot(raw, aes(x = iteration, colour = prof_covers)) +
  
    geom_linerange(aes(ymin = prof_lower, ymax = prof_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = prof_rl), shape = 16, size = 1.5) +
  
    scale_colour_manual(
      name   = "Covers True Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor  = element_blank(),
      plot.title        = element_text(face = "bold", hjust = 0.5),
      legend.position   = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin     = margin(4, 6, 4, 6)
    )
  
  list(
    delta   = p_delta,
    profile = p_prof
  )
}

```

## No seasonality run
```{r no seasonality case, eval=FALSE}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0, 12)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
nseg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
noseasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
noseasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
noseasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(noseasonality25_gumbel, 
     noseasonality50_gumbel, 
     noseasonality100_gumbel,  
     nseg_gumbel,
     file = "no_seasonality_gumbel.RData")

```

```{r No seasonality Data}
load("no_seasonality_gumbel.RData")

noseasonality_gumbel <- rbind(
  noseasonality25_gumbel$summary,
  noseasonality50_gumbel$summary,
  noseasonality100_gumbel$summary
)

noseasonality_gumbel
plot(nseg_gumbel[1:(5*365)])
nsbm_gumbel <- block_max(nseg_gumbel, 365)
nsfit_gumbel <- gev.fit(nsbm_gumbel)
gev.diag(nsfit_gumbel)

nspfit_gumbel <- gev_mle(nsbm_gumbel,1)
plot(nspfit_gumbel, profile = TRUE)

plots <- coverage_plots(
  noseasonality100_gumbel$raw, 
  noseasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile

```


```{r no seasonality for latex}
load("no_seasonality_gumbel.RData")

noseasonality_gumbel <- rbind(
  noseasonality25_gumbel$summary,
  noseasonality50_gumbel$summary,
  noseasonality100_gumbel$summary
)

################################################################################

estimation_table_delta <- noseasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    delta_rl,
    RMSE_delta,
    Bias_delta,
    Bias_delta_med,
    Coverage_delta,
    delta_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Delta RL" = delta_rl,
    "RMSE" = RMSE_delta,
    "Bias" = Bias_delta,
    "MD Bias" = Bias_delta_med,
    "Coverage" = Coverage_delta,
    "Width" = delta_width
    
  )

estimation_table_delta <- estimation_table_delta |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_delta |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta Return level estimation performance of no seasonality Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("noseasonality_gumbel_delta_rl.tex")

###############################################################################

estimation_table_prof <- noseasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    prof_rl,
    RMSE_prof,
    Bias_prof,
    Bias_prof_med,
    Coverage_prof,
    prof_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Profile RL" = prof_rl,
    "RMSE" = RMSE_prof,
    "Bias" = Bias_prof,
    "MD Bias" = Bias_prof_med,
    "Coverage" = Coverage_prof,
    "Width" = prof_width
  )

estimation_table_prof <- estimation_table_prof |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_prof |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood Return level estimation performance of no seasonality Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("noseasonality_gumbel_prof_rl.tex")

```

## S1 in paper and how to save the table of results
```{r version 1 seasonal case}
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v1seg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v1seasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v1seasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v1seasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v1seasonality25_gumbel, 
     v1seasonality50_gumbel, 
     v1seasonality100_gumbel,  
     v1seg_gumbel,
     file = "v1_seasonality_gumbel.RData")
```

```{r Plots and v1 seasonal Data}
load("v1_seasonality_gumbel.RData")

#v1seasonality25_gumbel
#v1seasonality50_gumbel
#v1seasonality100_gumbel

v1seasonality_gumbel <- rbind(
  v1seasonality25_gumbel$summary,
  v1seasonality50_gumbel$summary,
  v1seasonality100_gumbel$summary
)

v1seasonality_gumbel
plot(v1seg_gumbel[1:(5*365)])
v1sbm_gumbel <- block_max(v1seg_gumbel, 365)
v1sfit_gumbel <- gev.fit(v1sbm_gumbel)
gev.diag(v1sfit_gumbel)

plots <- coverage_plots(
  v1seasonality100_gumbel$raw, 
  v1seasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile

v1seasonality_gumbel



```

```{r Saving the table of results in latex}
load("v1_seasonality_gumbel.RData")

v1seasonality_gumbel <- rbind(
  v1seasonality25_gumbel$summary,
  v1seasonality50_gumbel$summary,
  v1seasonality100_gumbel$summary
)


################################################################################

estimation_table_delta <- v1seasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    delta_rl,
    RMSE_delta,
    Bias_delta,
    Bias_delta_med,
    Coverage_delta,
    delta_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Delta RL" = delta_rl,
    "RMSE" = RMSE_delta,
    "Bias" = Bias_delta,
    "MD Bias" = Bias_delta_med,
    "Coverage" = Coverage_delta,
    "Width" = delta_width
    
  )

estimation_table_delta <- estimation_table_delta |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_delta |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta Return level estimation performance of seasonality version 1 Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("v1seasonality_gumbel_delta_rl.tex")

###############################################################################

estimation_table_prof <- v1seasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    prof_rl,
    RMSE_prof,
    Bias_prof,
    Bias_prof_med,
    Coverage_prof,
    prof_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Profile RL" = prof_rl,
    "RMSE" = RMSE_prof,
    "Bias" = Bias_prof,
    "MD Bias" = Bias_prof_med,
    "Coverage" = Coverage_prof,
    "Width" = prof_width
  )

estimation_table_prof <- estimation_table_prof |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_prof |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood Return level estimation performance of seasonality version 1 Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("v1seasonality_gumbel_prof_rl.tex")
```

## S2 in paper and how to save the table of results
```{r version 2 seasonal case}
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,4,8,13,20,20,13,8,4,2,1)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v2seg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v2seasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v2seasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v2seasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v2seasonality25_gumbel, 
     v2seasonality50_gumbel, 
     v2seasonality100_gumbel,  
     v2seg_gumbel,
     file = "v2_seasonality_gumbel.RData")
```

```{r V2 Seasonality Data}
load("v2_seasonality_gumbel.RData")

#v2seasonality25_gumbel
#v2seasonality50_gumbel
#v2seasonality100_gumbel

v2seasonality_gumbel <- rbind(
  v2seasonality25_gumbel$summary,
  v2seasonality50_gumbel$summary,
  v2seasonality100_gumbel$summary
)

v2seasonality_gumbel
plot(v2seg_gumbel[1:(5*365)])
v2sbm_gumbel <- block_max(v2seg_gumbel, 365)
v2sfit_gumbel <- gev.fit(v2sbm_gumbel)
gev.diag(v2sfit_gumbel)

plots <- coverage_plots(
  v2seasonality100_gumbel$raw, 
  v2seasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile
```

```{r Saving the table of results in latex}
load("v2_seasonality_gumbel.RData")

v2seasonality_gumbel <- rbind(
  v2seasonality25_gumbel$summary,
  v2seasonality50_gumbel$summary,
  v2seasonality100_gumbel$summary
)

################################################################################

estimation_table_delta <- v2seasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    delta_rl,
    RMSE_delta,
    Bias_delta,
    Bias_delta_med,
    Coverage_delta,
    delta_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Delta RL" = delta_rl,
    "RMSE" = RMSE_delta,
    "Bias" = Bias_delta,
    "MD Bias" = Bias_delta_med,
    "Coverage" = Coverage_delta,
    "Width" = delta_width
    
  )

estimation_table_delta <- estimation_table_delta |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_delta |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta Return level estimation performance of seasonality version 2 Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("v2seasonality_gumbel_delta_rl.tex")



###############################################################################

estimation_table_prof <- v2seasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    prof_rl,
    RMSE_prof,
    Bias_prof,
    Bias_prof_med,
    Coverage_prof,
    prof_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Profile RL" = prof_rl,
    "RMSE" = RMSE_prof,
    "Bias" = Bias_prof,
    "MD Bias" = Bias_prof_med,
    "Coverage" = Coverage_prof,
    "Width" = prof_width
  )

estimation_table_prof <- estimation_table_prof |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_prof |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood Return level estimation performance of seasonality version 2 Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("v2seasonality_gumbel_prof_rl.tex")


```

## S4 in paper and how to save the table of results
WARNING: I realised later that I did not include an isolated case of keeping mean constant while standard deviation is varied so this is put to S4 and the other is S3
```{r version 3 seasonal case}
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- c(1,2,3,4,5,6,6,5,4,3,2,1)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v3seg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v3seasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v3seasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v3seasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v3seasonality25_gumbel, 
     v3seasonality50_gumbel, 
     v3seasonality100_gumbel,  
     v3seg_gumbel,
     file = "v3_seasonality_gumbel.RData")
```

```{r V3 Seasonality Data}
load("v3_seasonality_gumbel.RData")

#v3seasonality25_gumbel
#v3seasonality50_gumbel
#v3seasonality100_gumbel

v3seasonality_gumbel <- rbind(
  v3seasonality25_gumbel$summary,
  v3seasonality50_gumbel$summary,
  v3seasonality100_gumbel$summary
)

v3seasonality_gumbel
plot(v3seg_gumbel[1:(5*365)])
v3sbm_gumbel <- block_max(v3seg_gumbel, 365)
v3sfit_gumbel <- gev.fit(v3sbm_gumbel)
gev.diag(v3sfit_gumbel)

plots <- coverage_plots(
  v3seasonality100_gumbel$raw, 
  v3seasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile
```

```{r Saving the table of results in latex}
load("v3_seasonality_gumbel.RData")

v3seasonality_gumbel <- rbind(
  v3seasonality25_gumbel$summary,
  v3seasonality50_gumbel$summary,
  v3seasonality100_gumbel$summary
)

################################################################################

estimation_table_delta <- v3seasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    delta_rl,
    RMSE_delta,
    Bias_delta,
    Bias_delta_med,
    Coverage_delta,
    delta_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Delta RL" = delta_rl,
    "RMSE" = RMSE_delta,
    "Bias" = Bias_delta,
    "MD Bias" = Bias_delta_med,
    "Coverage" = Coverage_delta,
    "Width" = delta_width
    
  )

estimation_table_delta <- estimation_table_delta |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_delta |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta Return level estimation performance of seasonality version 3 Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("v3seasonality_gumbel_delta_rl.tex")

###############################################################################

estimation_table_prof <- v3seasonality_gumbel |>
  select(
    ReturnPeriod,
    true_rl,
    prof_rl,
    RMSE_prof,
    Bias_prof,
    Bias_prof_med,
    Coverage_prof,
    prof_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Profile RL" = prof_rl,
    "RMSE" = RMSE_prof,
    "Bias" = Bias_prof,
    "MD Bias" = Bias_prof_med,
    "Coverage" = Coverage_prof,
    "Width" = prof_width
  )

estimation_table_prof <- estimation_table_prof |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_prof |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood Return level estimation performance of seasonality version 3 Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("v3seasonality_gumbel_prof_rl.tex")


```

# FAT TABLE HERE THAT IS GOING TO BE COMBINED FOR THE LATEX CODE 
```{r Combination of the tables to make one fat one that looks better}

load("no_seasonality_gumbel.RData")

noseasonality_gumbel <- rbind(
  noseasonality25_gumbel$summary,
  noseasonality50_gumbel$summary,
  noseasonality100_gumbel$summary
)

load("v1_seasonality_gumbel.RData")

v1seasonality_gumbel <- rbind(
  v1seasonality25_gumbel$summary,
  v1seasonality50_gumbel$summary,
  v1seasonality100_gumbel$summary
)

load("v2_seasonality_gumbel.RData")

v2seasonality_gumbel <- rbind(
  v2seasonality25_gumbel$summary,
  v2seasonality50_gumbel$summary,
  v2seasonality100_gumbel$summary
)

load("v3_seasonality_gumbel.RData")

v3seasonality_gumbel <- rbind(
  v3seasonality25_gumbel$summary,
  v3seasonality50_gumbel$summary,
  v3seasonality100_gumbel$summary
)

# Flip this and v3 as I need v4 to be the important one as well

load("v4_seasonality_gumbel.RData")

v4seasonality_gumbel <- rbind(
  v4seasonality25_gumbel$summary,
  v4seasonality50_gumbel$summary,
  v4seasonality100_gumbel$summary
)

prepare_table_delta <- function(x, scenario, method){
  
  if(method == "Delta"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        delta_rl,
        RMSE_delta,
        Bias_delta,
        Bias_delta_med,
        Coverage_delta,
        delta_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = delta_rl,
        RMSE = RMSE_delta,
        Bias = Bias_delta,
        `MD Bias` = Bias_delta_med,
        Coverage = Coverage_delta,
        Width = delta_width
      )
    
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}



prepare_table_prof <- function(x, scenario, method){
  
  if(method == "Profile"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        prof_rl,
        RMSE_prof,
        Bias_prof,
        Bias_prof_med,
        Coverage_prof,
        prof_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = prof_rl,
        RMSE = RMSE_prof,
        Bias = Bias_prof,
        `MD Bias` = Bias_prof_med,
        Coverage = Coverage_prof,
        Width = prof_width
      )
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}


delta_all <- rbind(
  prepare_table_delta(noseasonality_gumbel, "NS", "Delta"),
  prepare_table_delta(v1seasonality_gumbel, "S1", "Delta"),
  prepare_table_delta(v2seasonality_gumbel, "S2", "Delta"),
  prepare_table_delta(v4seasonality_gumbel, "S3", "Delta"),
  prepare_table_delta(v3seasonality_gumbel, "S4", "Delta")
)

prof_all <- rbind(
  prepare_table_prof(noseasonality_gumbel, "NS", "Profile"),
  prepare_table_prof(v1seasonality_gumbel, "S1", "Profile"),
  prepare_table_prof(v2seasonality_gumbel, "S2", "Profile"),
  prepare_table_prof(v4seasonality_gumbel, "S3", "Profile"),
  prepare_table_prof(v3seasonality_gumbel, "S4", "Profile")
)

delta_all <- delta_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

prof_all <- prof_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )


delta_all
prof_all


delta_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta method for Return level estimation performance for Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("all_seasonality_gumbel_delta.tex")

prof_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood method for Return level estimation performance for Gumbel.",
    label = "tab:estimation"
  ) |>
  save_kable("all_seasonality_gumbel_prof.tex")





```

# Standardised Fat table when I realised that I cannot just compare the values like that 
```{r Gumbel Standardised table}

load("no_seasonality_gumbel.RData")

noseasonality_gumbel <- rbind(
  noseasonality25_gumbel$summary,
  noseasonality50_gumbel$summary,
  noseasonality100_gumbel$summary
)

load("v1_seasonality_gumbel.RData")

v1seasonality_gumbel <- rbind(
  v1seasonality25_gumbel$summary,
  v1seasonality50_gumbel$summary,
  v1seasonality100_gumbel$summary
)

load("v2_seasonality_gumbel.RData")

v2seasonality_gumbel <- rbind(
  v2seasonality25_gumbel$summary,
  v2seasonality50_gumbel$summary,
  v2seasonality100_gumbel$summary
)

load("v3_seasonality_gumbel.RData")

v3seasonality_gumbel <- rbind(
  v3seasonality25_gumbel$summary,
  v3seasonality50_gumbel$summary,
  v3seasonality100_gumbel$summary
)

# Flip this and v3 as I need v4 to be the important one as well

load("v4_seasonality_gumbel.RData")

v4seasonality_gumbel <- rbind(
  v4seasonality25_gumbel$summary,
  v4seasonality50_gumbel$summary,
  v4seasonality100_gumbel$summary
)


LS_gumbel <- tibble::tribble(
  ~Scenario, ~L, ~S,
  "NS",       0,  1,
  "S1",       6,  1,
  "S2",      20,  1,
  "S3",       0,  6,
  "S4",       6,  6
)

prepare_combined <- function(x, scenario) {
  x |>
    select(
      ReturnPeriod,
      true_rl,
      prof_rl, delta_rl,
      RMSE_prof, RMSE_delta,
      Bias_prof, Bias_prof_med,
      Coverage_delta, delta_width,
      Coverage_prof,  prof_width
    ) |>
    mutate(Scenario = scenario, .before = 1)
}

combined_raw <- rbind(
  prepare_combined(noseasonality_gumbel, "NS"),
  prepare_combined(v1seasonality_gumbel, "S1"),
  prepare_combined(v2seasonality_gumbel, "S2"),
  prepare_combined(v4seasonality_gumbel, "S3"),   # v4 -> S3
  prepare_combined(v3seasonality_gumbel, "S4")    # v3 -> S4
)

# lil checker incase the point estimates dont line up 
max_diff <- max(abs(combined_raw$prof_rl   - combined_raw$delta_rl),
                abs(combined_raw$RMSE_prof - combined_raw$RMSE_delta))
if (max_diff > 5e-4) {
  warning("Delta and profile point estimates differ by up to ",
          round(max_diff, 4),
          ". Table reports the profile values only -- check before submitting.")
}

build_gumbel_table <- function(dat, LS, standardise = TRUE) {

  dat |>
    left_join(LS, by = "Scenario") |>
    mutate(
      shift = if (standardise) L else 0,
      scale = if (standardise) S else 1
    ) |>
    transmute(
      Scenario,
      T        = ReturnPeriod,
      Ref      = (true_rl - shift) / scale,
      Est      = (prof_rl - shift) / scale,
      RMSE     = RMSE_prof     / scale,
      Bias     = Bias_prof     / scale,
      `M Bias` = Bias_prof_med / scale,
      DCovers  = Coverage_delta,
      DWidth   = delta_width   / scale,
      PCovers  = Coverage_prof,
      PWidth   = prof_width    / scale
    )
}

print_gumbel_table <- function(tab, caption, label, file,
                               digits = 3,
                               digits_by_col = NULL) {

  num_cols <- names(tab)[vapply(tab, is.numeric, logical(1))]
  num_cols <- setdiff(num_cols, "T")

  dg <- setNames(rep(digits, length(num_cols)), num_cols)
  if (!is.null(digits_by_col)) {
    dg[names(digits_by_col)] <- digits_by_col
  }

  chr <- tab
  for (nm in num_cols) {
    chr[[nm]] <- sprintf(paste0("%.", dg[[nm]], "f"), chr[[nm]])
  }
  chr$T <- as.character(chr$T)

  chr$Scenario[duplicated(chr$Scenario)] <- ""

  grp <- "& & \\multicolumn{2}{c}{Return Level} & \\multicolumn{3}{c}{Error} & \\multicolumn{2}{c}{Delta} & \\multicolumn{2}{c}{Profile} \\\\"
  mid <- "\\cmidrule(lr){3-4} \\cmidrule(lr){5-7} \\cmidrule(lr){8-9} \\cmidrule(lr){10-11}"
  nms <- "Scenario & T & Ref & Est & RMSE & Bias & M Bias & Covers & Width & Covers & Width \\\\"

  header <- paste("\\toprule", grp, mid, nms, "\\midrule", sep = "\n")

  n_per  <- 3L
  breaks <- seq(n_per, nrow(chr) - n_per, by = n_per)

  addtorow <- list(
    pos     = c(list(-1), as.list(breaks), list(nrow(chr))),
    command = c(header, rep("\\midrule\n", length(breaks)), "\\bottomrule\n")
  )

  xt <- xtable(chr,
               caption = caption,
               label   = label,
               align   = paste0("ll", strrep("r", ncol(chr) - 1)))

  print(xt,
        file              = file,
        include.rownames  = FALSE,
        include.colnames  = FALSE,
        hline.after       = NULL,
        add.to.row        = addtorow,
        booktabs          = TRUE,
        caption.placement = "top",
        sanitize.text.function = identity,
        table.placement   = "ht")
}

std_table <- build_gumbel_table(combined_raw, LS_gumbel, standardise = TRUE)

print_gumbel_table(
  std_table,
  caption = "...",
  label   = "tab:gumbel_standardised",
  file    = "gumbel_standardised.tex",
  digits  = 3,
  digits_by_col = c(Ref = 3, Est = 3, DCovers = 3, PCovers = 3)
)

std_table

```

# Attempt at generating boxplots for xi here what can I do? 
```{r Generating boxplots for xi}

load("no_seasonality_gumbel.RData")

noseasonality_gumbel <- rbind(
  noseasonality25_gumbel$summary,
  noseasonality50_gumbel$summary,
  noseasonality100_gumbel$summary
)

load("v1_seasonality_gumbel.RData")

v1seasonality_gumbel <- rbind(
  v1seasonality25_gumbel$summary,
  v1seasonality50_gumbel$summary,
  v1seasonality100_gumbel$summary
)

load("v2_seasonality_gumbel.RData")

v2seasonality_gumbel <- rbind(
  v2seasonality25_gumbel$summary,
  v2seasonality50_gumbel$summary,
  v2seasonality100_gumbel$summary
)

load("v3_seasonality_gumbel.RData")

v3seasonality_gumbel <- rbind(
  v3seasonality25_gumbel$summary,
  v3seasonality50_gumbel$summary,
  v3seasonality100_gumbel$summary
)

# Flip this and v3 as I need v4 to be the important one as well

load("v4_seasonality_gumbel.RData")

v4seasonality_gumbel <- rbind(
  v4seasonality25_gumbel$summary,
  v4seasonality50_gumbel$summary,
  v4seasonality100_gumbel$summary
)

noseasonality100_gumbel$
```

# Version for S3 test case
```{r version 4 seasonal case}
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0,12)
sigma       <- c(1,2,3,4,5,6,6,5,4,3,2,1)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v4seg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v4seasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v4seasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v4seasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v4seasonality25_gumbel, 
     v4seasonality50_gumbel, 
     v4seasonality100_gumbel,  
     v4seg_gumbel,
     file = "v4_seasonality_gumbel.RData")
```

```{r V4 Seasonality Data}
load("v4_seasonality_gumbel.RData")

#v3seasonality25_gumbel
#v3seasonality50_gumbel
#v3seasonality100_gumbel

v4seasonality_gumbel <- rbind(
  v4seasonality25_gumbel$summary,
  v4seasonality50_gumbel$summary,
  v4seasonality100_gumbel$summary
)

v4seasonality_gumbel
plot(v4seg_gumbel[1:(5*365)])
v4sbm_gumbel <- block_max(v4seg_gumbel, 365)
v4sfit_gumbel <- gev.fit(v4sbm_gumbel)
gev.diag(v4sfit_gumbel)

plots <- coverage_plots(
  v4seasonality100_gumbel$raw, 
  v4seasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile
```

# 16 year case comparison
```{r V4 Thinner seasonality}

GEV_samples <- 16 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0, 12)
sigma       <- rep(1,12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v4seg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v4seasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v4seasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v4seasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v4seasonality25_gumbel, 
     v4seasonality50_gumbel, 
     v4seasonality100_gumbel,  
     v4seg_gumbel,
     file = "v4_seasonality_gumbel.RData")
```

```{r V4 seasonality Data}

load("v4_seasonality_gumbel.RData")

v4seasonality_gumbel <- rbind(
  v4seasonality25_gumbel$summary,
  v4seasonality50_gumbel$summary,
  v4seasonality100_gumbel$summary
)

v4seasonality100_gumbel
v4seasonality_gumbel
plot(v4seg_gumbel[1:(5*365)])
v4sbm_gumbel <- block_max(v4seg_gumbel, 365)
v4sfit_gumbel <- gev.fit(v4sbm_gumbel)
gev.diag(v4sfit_gumbel)

plots <- coverage_plots(
  v4seasonality100_gumbel$raw, 
  v4seasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile
  
```

# Test case for another seasonality version which I should look into 
```{r V5 Seasonal implementation}
GEV_samples <- 16 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- rep(1,12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v5seg_gumbel <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v5seasonality25_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v5seasonality50_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v5seasonality100_gumbel <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v5seasonality25_gumbel, 
     v5seasonality50_gumbel, 
     v5seasonality100_gumbel,  
     v5seg_gumbel,
     file = "v5_seasonality_gumbel.RData")
```
Note that I think the 48th row one is the absolutely disgustingly big return level for some reason 
Could look to remove this point 

```{r V5 Seasonal Data}
load("v5_seasonality_gumbel.RData")

v5seasonality_gumbel <- rbind(
  v5seasonality25_gumbel$summary,
  v5seasonality50_gumbel$summary,
  v5seasonality100_gumbel$summary
)

v5seasonality_gumbel
plot(v5seg_gumbel[1:(5*365)])
v5sbm_gumbel <- block_max(v5seg_gumbel, 365)
v5sfit_gumbel <- gev.fit(v5sbm_gumbel)
gev.diag(v5sfit_gumbel)

v5seasonality100_gumbel

plots <- coverage_plots(
  v5seasonality100_gumbel$raw, 
  v5seasonality100_gumbel$summary$true_rl
  )

plots$delta
plots$profile


```




