---
title: "Frechet Simulations"
author: '25062476'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ismev) # Coles Package
library(evmissing) # Simpson & Northrop package 
library(readr) # for csv and things 
library(ggplot2)


library(dplyr)
library(knitr)
library(kableExtra)
```

# This is generally used for Chapter 3 / Appendix for the Frechet domain
Same idea as the Gumbel case

```{r Frechet big function}
x_samples <- function(n_samples, mu, sigma, df = 10) {

  month_days <- c(31,28,31,30,31,30,31,31,30,31,30,31)

  month_index <- rep(1:12, times = month_days)
  month_index <- rep(month_index, n_samples / 365)

  mu[month_index] +
    sigma[month_index] * rt(length(month_index), df = df)
}

df <- 10

block_max <- function(x, block_size) {
  as.numeric(tapply(x, (seq_along(x) - 1) %/% block_size, max))
}

delta_rl_CI <- function(T, fit) {
  mu    <- fit$mle[1]
  sigma <- fit$mle[2]
  xi    <- fit$mle[3]
  V     <- fit$cov
  y     <- -log(1 - 1/T)

  if (abs(xi) < 1e-6) {
    rl     <- mu - sigma * log(y)
    var_rl <- V[1,1] + (log(y))^2 * V[2,2] - 2 * log(y) * V[1,2]
  } else {
    rl      <- mu + (sigma/xi) * (y^(-xi) - 1)
    grad_rl <- c(1, (y^(-xi) - 1)/xi, sigma * (1 - y^(-xi) - xi * y^(-xi) * log(y))/xi^2)
    var_rl  <- as.numeric(t(grad_rl) %*% V %*% grad_rl)
  }

  se_rl <- sqrt(var_rl)
  data.frame(ReturnLevel = rl, Lower = rl - 1.96 * se_rl, Upper = rl + 1.96 * se_rl)
}

delta_prof <- function(GEV_samples, n_samples, q_samples, T, mu, sigma) {

  # True return level from large sample
  x_true  <- x_samples(q_samples, mu, sigma)
  bm_true <- block_max(x_true, 365)
  true_rl <- as.numeric(quantile(bm_true, probs = 1 - 1/T))

  cat("True return level (T =", T, "):", round(true_rl, 4), "\n")
  
  results <- data.frame(
    delta_rl    = numeric(n_samples),
    delta_lower = numeric(n_samples),
    delta_upper = numeric(n_samples),
    prof_rl     = numeric(n_samples),
    prof_lower  = numeric(n_samples),
    prof_upper  = numeric(n_samples)
  )

  delta_fail     <- 0L
  profile_fail   <- 0L
  profileci_fail <- 0L

  for (i in seq_len(n_samples)) {

    if (i %% max(1, n_samples %/% 4) == 0)
      cat(round(100 * i / n_samples), "% done\n")

    x_dummy  <- x_samples(GEV_samples, mu, sigma)
    bm_dummy <- block_max(x_dummy, 365)

    # Delta method
    fit_delta <- tryCatch(
      suppressWarnings(gev.fit(bm_dummy, show = FALSE)),
      error = function(e) NULL
    )

    if (is.null(fit_delta)) {
      delta_fail <- delta_fail + 1L
      results[i, c("delta_rl", "delta_lower", "delta_upper")] <- NA
    } else {
      ci <- delta_rl_CI(T, fit_delta)
      results[i, "delta_rl"]    <- ci$ReturnLevel
      results[i, "delta_lower"] <- ci$Lower
      results[i, "delta_upper"] <- ci$Upper
    }

    # Profile likelihood
    fit_prof <- tryCatch(
      suppressWarnings(gev_mle(bm_dummy, 1)),
      error = function(e) NULL
    )

    if (is.null(fit_prof) || any(is.na(fit_prof))) {
      profile_fail <- profile_fail + 1L
      results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
    } else {
      prof_return <- gev_return(fit_prof, m = T)
      prof_conf   <- tryCatch(
        suppressWarnings(confint(prof_return, profile = TRUE)),
        error = function(e) NULL
      )

      if (is.null(prof_conf)) {
        profileci_fail <- profileci_fail + 1L
        results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
      } else {
        results[i, "prof_rl"]    <- as.numeric(prof_return)
        results[i, "prof_lower"] <- prof_conf[1, 1]
        results[i, "prof_upper"] <- prof_conf[1, 2]
      }
    }
  }

  # Summary
  summary_df <- data.frame(
    ReturnPeriod = T,
    true_rl      = true_rl,

    delta_rl   = mean(results$delta_rl,  na.rm = TRUE),
    prof_rl    = mean(results$prof_rl,   na.rm = TRUE),

    RMSE_delta = sqrt(mean((results$delta_rl - true_rl)^2, na.rm = TRUE)),
    RMSE_prof  = sqrt(mean((results$prof_rl  - true_rl)^2, na.rm = TRUE)),

    Bias_delta     = mean(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof      = mean(results$prof_rl  - true_rl, na.rm = TRUE),

    Bias_delta_med = median(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof_med  = median(results$prof_rl  - true_rl, na.rm = TRUE),

    Coverage_delta = mean(results$delta_lower <= true_rl &
                          results$delta_upper >= true_rl, na.rm = TRUE),
    Coverage_prof  = mean(results$prof_lower  <= true_rl &
                          results$prof_upper  >= true_rl, na.rm = TRUE),
    
    delta_width    = mean(results$delta_upper - results$delta_lower, na.rm=TRUE),
    prof_width     = mean(results$prof_upper - results$prof_lower, na.rm=TRUE)
    
  )

  list(
    summary            = summary_df,
    raw                = results,        
    true_rl            = true_rl,
    delta_failures     = delta_fail,
    profile_failures   = profile_fail,
    profileci_failures = profileci_fail,
    delta_failure_rate   = delta_fail     / n_samples,
    profile_failure_rate = profile_fail   / n_samples
  )
}


coverage_plots <- function(raw, true_rl) {
  
  raw     <- raw
  true_rl <- true_rl
  
  n <- nrow(raw)
  raw$iteration <- seq_len(n)
  
  raw$delta_covers <- raw$delta_lower <= true_rl & raw$delta_upper >= true_rl
  raw$prof_covers  <- raw$prof_lower  <= true_rl & raw$prof_upper  >= true_rl
  
  # ── Delta plot ────────────────────────────────────────────────────────────────
  p_delta <- ggplot(raw, aes(x = iteration, colour = delta_covers)) +
  
    geom_linerange(aes(ymin = delta_lower, ymax = delta_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = delta_rl), shape = 16, size = 1.5, alpha = 0.5) +
  
    scale_colour_manual(
      name   = "Covers True Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", hjust = 0.5),
      legend.position  = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin    = margin(4, 6, 4, 6)
    )
  
  # ── Profile plot ──────────────────────────────────────────────────────────────
  p_prof <- ggplot(raw, aes(x = iteration, colour = prof_covers)) +
  
    geom_linerange(aes(ymin = prof_lower, ymax = prof_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = prof_rl), shape = 16, size = 1.5) +
  
    scale_colour_manual(
      name   = "Covers True Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor  = element_blank(),
      plot.title        = element_text(face = "bold", hjust = 0.5),
      legend.position   = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin     = margin(4, 6, 4, 6)
    )
  
  list(
    delta   = p_delta,
    profile = p_prof
  )
}
```


## This is the noseasonality case for frechet into no_seasonality_frechet
NOTE WE SHOULD CHANGE THE INSIDE CODE FOR x_samples AS REQUIRED SO RE RUN THE CODE ABOVE 
Also rather than making individual tables for all versions here, we combine them at the bottom for one fat table 

```{r No seasonality}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0, 12)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
nseg_frechet <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
noseasonality25_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
noseasonality50_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
noseasonality100_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(noseasonality25_frechet, 
     noseasonality50_frechet, 
     noseasonality100_frechet,  
     nseg_frechet,
     file = "no_seasonality_frechet.RData")
```

```{r No seasonality Data}
load("no_seasonality_frechet.RData")

noseasonality_frechet <- rbind(
  noseasonality25_frechet$summary,
  noseasonality50_frechet$summary,
  noseasonality100_frechet$summary
)

noseasonality_frechet
plot(nseg_frechet[1:(5*365)])
nsbm_frechet <- block_max(nseg_frechet, 365)
nsfit_frechet <- gev.fit(nsbm_frechet)
gev.diag(nsfit_frechet)

noseasonality100_frechet

plots <- coverage_plots(
  noseasonality100_frechet$raw, 
  noseasonality100_frechet$summary$true_rl
  )

plots$delta
plots$profile


```

```{r No seasonal data table generation}

load("no_seasonality_frechet.RData")

noseasonality_frechet <- rbind(
  noseasonality25_frechet$summary,
  noseasonality50_frechet$summary,
  noseasonality100_frechet$summary
)

################################################################################

estimation_table_delta <- noseasonality_frechet |>
  select(
    ReturnPeriod,
    true_rl,
    delta_rl,
    RMSE_delta,
    Bias_delta,
    Bias_delta_med,
    Coverage_delta,
    delta_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Delta RL" = delta_rl,
    "RMSE" = RMSE_delta,
    "Bias" = Bias_delta,
    "MD Bias" = Bias_delta_med,
    "Coverage" = Coverage_delta,
    "Width" = delta_width
    
  )

estimation_table_delta <- estimation_table_delta |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_delta |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta Return level estimation performance of no seasonality Frechet.",
    label = "tab:estimation"
  ) |>
  save_kable("noseasonality_frechet_delta_rl.tex")

###############################################################################

estimation_table_prof <- noseasonality_frechet |>
  select(
    ReturnPeriod,
    true_rl,
    prof_rl,
    RMSE_prof,
    Bias_prof,
    Bias_prof_med,
    Coverage_prof,
    prof_width
  ) |>
  rename(
    "T" = ReturnPeriod,
    "True RL" = true_rl,
    "Profile RL" = prof_rl,
    "RMSE" = RMSE_prof,
    "Bias" = Bias_prof,
    "MD Bias" = Bias_prof_med,
    "Coverage" = Coverage_prof,
    "Width" = prof_width
  )

estimation_table_prof <- estimation_table_prof |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

estimation_table_prof |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood Return level estimation performance of no seasonality Frechet.",
    label = "tab:estimation"
  ) |>
  save_kable("noseasonality_frechet_prof_rl.tex")

```


# S1 
```{r Seasonal V1}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v1seg_frechet <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v1seasonality25_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v1seasonality50_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v1seasonality100_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v1seasonality25_frechet, 
     v1seasonality50_frechet, 
     v1seasonality100_frechet,  
     v1seg_frechet,
     file = "v1_seasonality_frechet.RData")
```

```{r V1 Seasonal Data}

load("v1_seasonality_frechet.RData")

v1seasonality_frechet <- rbind(
  v1seasonality25_frechet$summary,
  v1seasonality50_frechet$summary,
  v1seasonality100_frechet$summary
)

v1seasonality_frechet
plot(v1seg_frechet[1:(5*365)])
v1sbm_frechet <- block_max(v1seg_frechet, 365)
v1sfit_frechet <- gev.fit(v1sbm_frechet)
gev.diag(v1sfit_frechet)

v1seasonality100_frechet

plots <- coverage_plots(
  v1seasonality100_frechet$raw, 
  v1seasonality100_frechet$summary$true_rl
  )

plots$delta
plots$profile

```

```{r For generating the table}
load("v1_seasonality_frechet.RData")

v1seasonality_frechet <- rbind(
  v1seasonality25_frechet$summary,
  v1seasonality50_frechet$summary,
  v1seasonality100_frechet$summary
)


```

# S2
```{r Seasonal V2}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,4,8,13,20,20,13,8,4,2,1)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v2seg_frechet <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v2seasonality25_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v2seasonality50_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v2seasonality100_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v2seasonality25_frechet, 
     v2seasonality50_frechet, 
     v2seasonality100_frechet,  
     v2seg_frechet,
     file = "v2_seasonality_frechet.RData")

```

```{r V2 seasonality Data}
load("v2_seasonality_frechet.RData")

v2seasonality_frechet <- rbind(
  v2seasonality25_frechet$summary,
  v2seasonality50_frechet$summary,
  v2seasonality100_frechet$summary
)

v2seasonality_frechet
plot(v2seg_frechet[1:(5*365)])
v2sbm_frechet <- block_max(v2seg_frechet, 365)
v2sfit_frechet <- gev.fit(v2sbm_frechet)
gev.diag(v2sfit_frechet)

v2seasonality100_frechet

plots <- coverage_plots(
  v2seasonality100_frechet$raw, 
  v2seasonality100_frechet$summary$true_rl
  )

plots$delta
plots$profile
```

# S4 as noted before in Gumbel that this changed
```{r Seasonal V3}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- c(1,2,3,4,5,6,6,5,4,3,2,1)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v3seg_frechet <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v3seasonality25_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v3seasonality50_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v3seasonality100_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v3seasonality25_frechet, 
     v3seasonality50_frechet, 
     v3seasonality100_frechet,  
     v3seg_frechet,
     file = "v3_seasonality_frechet.RData")

```

```{r V3 seasonal Data}

load("v3_seasonality_frechet.RData")

v3seasonality_frechet <- rbind(
  v3seasonality25_frechet$summary,
  v3seasonality50_frechet$summary,
  v3seasonality100_frechet$summary
)

v3seasonality_frechet
plot(v3seg_frechet[1:(5*365)])
v3sbm_frechet <- block_max(v3seg_frechet, 365)
v3sfit_frechet <- gev.fit(v3sbm_frechet)
gev.diag(v3sfit_frechet)

v3seasonality100_frechet

plots <- coverage_plots(
  v3seasonality100_frechet$raw, 
  v3seasonality100_frechet$summary$true_rl
  )

plots$delta
plots$profile
```

# S3
```{r Seasonal V4}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0,12)
sigma       <- c(1,2,3,4,5,6,6,5,4,3,2,1)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
v4seg_frechet <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
v4seasonality25_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
v4seasonality50_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
v4seasonality100_frechet <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(v4seasonality25_frechet, 
     v4seasonality50_frechet, 
     v4seasonality100_frechet,  
     v4seg_frechet,
     file = "v4_seasonality_frechet.RData")

```

```{r V4 seasonal Data}

load("v4_seasonality_frechet.RData")

v4seasonality_frechet <- rbind(
  v4seasonality25_frechet$summary,
  v4seasonality50_frechet$summary,
  v4seasonality100_frechet$summary
)

v4seasonality_frechet
plot(v4seg_frechet[1:(5*365)])
v4sbm_frechet <- block_max(v4seg_frechet, 365)
v4sfit_frechet <- gev.fit(v4sbm_frechet)
gev.diag(v4sfit_frechet)

v4seasonality100_frechet

plots <- coverage_plots(
  v4seasonality100_frechet$raw, 
  v4seasonality100_frechet$summary$true_rl
  )

plots$delta
plots$profile
```


# THIS IS THE BIG FAT TABLE
```{r Fat table}

load("no_seasonality_frechet.RData")

noseasonality_frechet <- rbind(
  noseasonality25_frechet$summary,
  noseasonality50_frechet$summary,
  noseasonality100_frechet$summary
)

load("v1_seasonality_frechet.RData")

v1seasonality_frechet <- rbind(
  v1seasonality25_frechet$summary,
  v1seasonality50_frechet$summary,
  v1seasonality100_frechet$summary
)

load("v2_seasonality_frechet.RData")

v2seasonality_frechet <- rbind(
  v2seasonality25_frechet$summary,
  v2seasonality50_frechet$summary,
  v2seasonality100_frechet$summary
)

load("v3_seasonality_frechet.RData")

v3seasonality_frechet <- rbind(
  v3seasonality25_frechet$summary,
  v3seasonality50_frechet$summary,
  v3seasonality100_frechet$summary
)

load("v4_seasonality_frechet.RData")

v4seasonality_frechet <- rbind(
  v4seasonality25_frechet$summary,
  v4seasonality50_frechet$summary,
  v4seasonality100_frechet$summary
)

prepare_table_delta <- function(x, scenario, method){
  
  if(method == "Delta"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        delta_rl,
        RMSE_delta,
        Bias_delta,
        Bias_delta_med,
        Coverage_delta,
        delta_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = delta_rl,
        RMSE = RMSE_delta,
        Bias = Bias_delta,
        `MD Bias` = Bias_delta_med,
        Coverage = Coverage_delta,
        Width = delta_width
      )
    
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}



prepare_table_prof <- function(x, scenario, method){
  
  if(method == "Profile"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        prof_rl,
        RMSE_prof,
        Bias_prof,
        Bias_prof_med,
        Coverage_prof,
        prof_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = prof_rl,
        RMSE = RMSE_prof,
        Bias = Bias_prof,
        `MD Bias` = Bias_prof_med,
        Coverage = Coverage_prof,
        Width = prof_width
      )
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}


delta_all <- rbind(
  prepare_table_delta(noseasonality_frechet, "NS", "Delta"),
  prepare_table_delta(v1seasonality_frechet, "S1", "Delta"),
  prepare_table_delta(v2seasonality_frechet, "S2", "Delta"),
  prepare_table_delta(v4seasonality_frechet, "S3", "Delta"),
  prepare_table_delta(v3seasonality_frechet, "S4", "Delta")
)

prof_all <- rbind(
  prepare_table_prof(noseasonality_frechet, "NS", "Profile"),
  prepare_table_prof(v1seasonality_frechet, "S1", "Profile"),
  prepare_table_prof(v2seasonality_frechet, "S2", "Profile"),
  prepare_table_prof(v4seasonality_frechet, "S3", "Profile"),
  prepare_table_prof(v3seasonality_frechet, "S4", "Profile")
)

delta_all <- delta_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

prof_all <- prof_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )


delta_all
prof_all


delta_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta method for Return level estimation performance for frechet.",
    label = "tab:estimation"
  ) |>
  save_kable("all_seasonality_frechet_delta.tex")

prof_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood method for Return level estimation performance for frechet.",
    label = "tab:estimation"
  ) |>
  save_kable("all_seasonality_frechet_prof.tex")





```

# Standardised version when I realised something is not right
```{r Standardised table}

load("no_seasonality_frechet.RData")

noseasonality_frechet <- rbind(
  noseasonality25_frechet$summary,
  noseasonality50_frechet$summary,
  noseasonality100_frechet$summary
)

load("v1_seasonality_frechet.RData")

v1seasonality_frechet <- rbind(
  v1seasonality25_frechet$summary,
  v1seasonality50_frechet$summary,
  v1seasonality100_frechet$summary
)

load("v2_seasonality_frechet.RData")

v2seasonality_frechet <- rbind(
  v2seasonality25_frechet$summary,
  v2seasonality50_frechet$summary,
  v2seasonality100_frechet$summary
)

load("v3_seasonality_frechet.RData")

v3seasonality_frechet <- rbind(
  v3seasonality25_frechet$summary,
  v3seasonality50_frechet$summary,
  v3seasonality100_frechet$summary
)

load("v4_seasonality_frechet.RData")

v4seasonality_frechet <- rbind(
  v4seasonality25_frechet$summary,
  v4seasonality50_frechet$summary,
  v4seasonality100_frechet$summary
)

LS_frechet <- tibble::tribble(
  ~Scenario, ~L, ~S,
  "NS",       0,  1,
  "S1",       6,  1,
  "S2",      20,  1,
  "S3",       0,  6,
  "S4",       6,  6
)

## ---------------------------------------------------------------
## 2. Pull delta and profile results into one row per scenario/T
## ---------------------------------------------------------------

prepare_combined <- function(x, scenario) {
  x |>
    select(
      ReturnPeriod,
      true_rl,
      prof_rl, delta_rl,
      RMSE_prof, RMSE_delta,
      Bias_prof, Bias_prof_med,
      Coverage_delta, delta_width,
      Coverage_prof,  prof_width
    ) |>
    mutate(Scenario = scenario, .before = 1)
}

combined_raw <- rbind(
  prepare_combined(noseasonality_frechet, "NS"),
  prepare_combined(v1seasonality_frechet, "S1"),
  prepare_combined(v2seasonality_frechet, "S2"),
  prepare_combined(v4seasonality_frechet, "S3"),   # v4 -> S3
  prepare_combined(v3seasonality_frechet, "S4")    # v3 -> S4
)

## Sanity check: point estimates should agree between methods
max_diff <- max(abs(combined_raw$prof_rl   - combined_raw$delta_rl),
                abs(combined_raw$RMSE_prof - combined_raw$RMSE_delta))
if (max_diff > 5e-4) {
  warning("Delta and profile point estimates differ by up to ",
          round(max_diff, 4),
          ". Table reports the profile values only -- check before submitting.")
}

## ---------------------------------------------------------------
## 3. Build the table, optionally standardised
## ---------------------------------------------------------------
build_frechet_table <- function(dat, LS, standardise = TRUE) {

  dat |>
    left_join(LS, by = "Scenario") |>
    mutate(
      shift = if (standardise) L else 0,
      scale = if (standardise) S else 1
    ) |>
    transmute(
      Scenario,
      T        = ReturnPeriod,
      Ref      = (true_rl - shift) / scale,
      Est      = (prof_rl - shift) / scale,
      RMSE     = RMSE_prof     / scale,
      Bias     = Bias_prof     / scale,
      `M Bias` = Bias_prof_med / scale,
      DCovers  = Coverage_delta,
      DWidth   = delta_width   / scale,
      PCovers  = Coverage_prof,
      PWidth   = prof_width    / scale
    )
}

print_frechet_table <- function(tab, caption, label, file,
                               digits = 3,
                               digits_by_col = NULL) {

  ## default: same precision everywhere
  num_cols <- names(tab)[vapply(tab, is.numeric, logical(1))]
  num_cols <- setdiff(num_cols, "T")

  dg <- setNames(rep(digits, length(num_cols)), num_cols)
  if (!is.null(digits_by_col)) {
    dg[names(digits_by_col)] <- digits_by_col
  }

  chr <- tab
  for (nm in num_cols) {
    chr[[nm]] <- sprintf(paste0("%.", dg[[nm]], "f"), chr[[nm]])
  }
  chr$T <- as.character(chr$T)

  chr$Scenario[duplicated(chr$Scenario)] <- ""

  grp <- "& & \\multicolumn{2}{c}{Return Level} & \\multicolumn{3}{c}{Error} & \\multicolumn{2}{c}{Delta} & \\multicolumn{2}{c}{Profile} \\\\"
  mid <- "\\cmidrule(lr){3-4} \\cmidrule(lr){5-7} \\cmidrule(lr){8-9} \\cmidrule(lr){10-11}"
  nms <- "Scenario & T & Ref & Est & RMSE & Bias & M Bias & Covers & Width & Covers & Width \\\\"

  header <- paste("\\toprule", grp, mid, nms, "\\midrule", sep = "\n")

  n_per  <- 3L
  breaks <- seq(n_per, nrow(chr) - n_per, by = n_per)

  addtorow <- list(
    pos     = c(list(-1), as.list(breaks), list(nrow(chr))),
    command = c(header, rep("\\midrule\n", length(breaks)), "\\bottomrule\n")
  )

  xt <- xtable(chr,
               caption = caption,
               label   = label,
               align   = paste0("ll", strrep("r", ncol(chr) - 1)))

  print(xt,
        file              = file,
        include.rownames  = FALSE,
        include.colnames  = FALSE,
        hline.after       = NULL,
        add.to.row        = addtorow,
        booktabs          = TRUE,
        caption.placement = "top",
        sanitize.text.function = identity,
        table.placement   = "ht")
}

std_table <- build_frechet_table(combined_raw, LS_frechet, standardise = TRUE)

print_frechet_table(
  std_table,
  caption = "...",
  label   = "tab:frechet_standardised",
  file    = "frechet_standardised.tex",
  digits  = 3,
  digits_by_col = c(Ref = 3, Est = 3, DCovers = 3, PCovers = 3)
)

std_table
```


# Interesting case of some interesting result 
Do not get rid of this one as this is probably one of the best odd cases you have 
From an older version of producing results 

```{r seasonal case 2, eval=FALSE}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
set.seed(123)

GEV_samples <- 100 * 365
n_samples <- 100 # number of n_samples 
q_samples <- 100000 * 365

rlcolnames <- c("rl10", "rl20", "rl100", "rl1000")
T <- c(10,20,100,1000)  

mu <- c(5,6,8,12,16,20,25,25,20,15,10,6)
sigma <- rep(2,12)

s2eg_frechet <- x_samples(GEV_samples, mu, sigma)
s2bm_frechet <- block_max(s2eg_frechet, 365)
s2fit_frechet <- gev.fit(s2bm_frechet)
s2diag_frechet <- gev.diag(s2fit_frechet)

set.seed(123)

seasonal2_frechet <- delta_prof(GEV_samples, n_samples, q_samples, rlcolnames, T, mu, sigma)

save(seasonal2_frechet, s2eg_frechet, s2bm_frechet, s2fit_frechet, s2diag_frechet, file = "seasonal2_frechet.RData")
# could remove the sdiag_frechet as it seems like it doesnt get saved correctly somehow and just apply the gev.diag after as it does not seem to take too long anyway

################################################################################
# To print the table from what we have saved 

load("seasonal2_frechet.RData")
s2fit_frechet
gev.diag(s2fit_frechet)
seasonal2_frechet
plot(s2eg_frechet[1:(5*365)])
plot(s2eg_frechet)
plot(s2bm_frechet)
```

```{r seasonal case 2, eval=FALSE}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
set.seed(123)

GEV_samples <- 16 * 365
n_samples <- 100 # number of n_samples 
q_samples <- 100000 * 365

rlcolnames <- c("rl10", "rl20", "rl100", "rl1000")
T <- c(10,20,100,1000)  

mu <- c(5,6,8,12,16,20,25,25,20,15,10,6)
sigma <- rep(2,12)

s3eg_frechet <- x_samples(GEV_samples, mu, sigma)
s3bm_frechet <- block_max(s3eg_frechet, 365)
s3fit_frechet <- gev.fit(s3bm_frechet)
s3diag_frechet <- gev.diag(s3fit_frechet)

set.seed(123)

seasonal3_frechet <- delta_prof(GEV_samples, n_samples, q_samples, rlcolnames, T, mu, sigma)

save(seasonal3_frechet, s3eg_frechet, s3bm_frechet, s3fit_frechet, s3diag_frechet, file = "seasonal3_frechet.RData")
# could remove the sdiag_frechet as it seems like it doesnt get saved correctly somehow and just apply the gev.diag after as it does not seem to take too long anyway

################################################################################
# To print the table from what we have saved 

load("seasonal3_frechet.RData")
s3fit_frechet
gev.diag(s3fit_frechet)
seasonal3_frechet
plot(s3eg_frechet[1:(5*365)])
plot(s3eg_frechet)
plot(s3bm_frechet)
```
