---
title: "Weibull Simulations"
author: '25062476'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ismev) # Coles Package
library(evmissing) # Simpson & Northrop package 
library(readr) # for csv and things 
library(ggplot2)

library(xtable)
library(dplyr)
library(knitr)
library(kableExtra)
```

# This is generally used for Chapter 3 / Appendix for the Weibull domain
Same idea as the Gumbel case

```{r Weibull big functions}

x_samples <- function(n_samples, mu, sigma, alpha = 1, beta = 10) {
  month_days  <- c(31,28,31,30,31,30,31,31,30,31,30,31)
  month_index <- rep(1:12, times = month_days)
  month_index <- rep(month_index, n_samples/365)

  mu[month_index] +
    sigma[month_index] * rbeta(length(month_index), alpha, beta)
}

block_max <- function(x, block_size) {
  as.numeric(tapply(x, (seq_along(x) - 1) %/% block_size, max))
}

delta_rl_CI <- function(T, fit) {
  mu    <- fit$mle[1]
  sigma <- fit$mle[2]
  xi    <- fit$mle[3]
  V     <- fit$cov
  y     <- -log(1 - 1/T)

  if (abs(xi) < 1e-6) {
    rl     <- mu - sigma * log(y)
    var_rl <- V[1,1] + (log(y))^2 * V[2,2] - 2 * log(y) * V[1,2]
  } else {
    rl      <- mu + (sigma/xi) * (y^(-xi) - 1)
    grad_rl <- c(1, (y^(-xi) - 1)/xi, sigma * (1 - y^(-xi) - xi * y^(-xi) * log(y))/xi^2)
    var_rl  <- as.numeric(t(grad_rl) %*% V %*% grad_rl)
  }

  se_rl <- sqrt(var_rl)
  data.frame(ReturnLevel = rl, Lower = rl - 1.96 * se_rl, Upper = rl + 1.96 * se_rl)
}

delta_prof <- function(GEV_samples, n_samples, q_samples, T, mu, sigma) {

  # True return level from large sample
  x_true  <- x_samples(q_samples, mu, sigma)
  bm_true <- block_max(x_true, 365)
  true_rl <- as.numeric(quantile(bm_true, probs = 1 - 1/T))

  cat("True return level (T =", T, "):", round(true_rl, 4), "\n")

  results <- data.frame(
    delta_rl    = numeric(n_samples),
    delta_lower = numeric(n_samples),
    delta_upper = numeric(n_samples),
    prof_rl     = numeric(n_samples),
    prof_lower  = numeric(n_samples),
    prof_upper  = numeric(n_samples)
  )

  delta_fail     <- 0L
  profile_fail   <- 0L
  profileci_fail <- 0L

  for (i in seq_len(n_samples)) {

    if (i %% max(1, n_samples %/% 4) == 0)
      cat(round(100 * i / n_samples), "% done\n")

    x_dummy  <- x_samples(GEV_samples, mu, sigma)
    bm_dummy <- block_max(x_dummy, 365)

    # Delta method
    fit_delta <- tryCatch(
      suppressWarnings(gev.fit(bm_dummy, show = FALSE)),
      error = function(e) NULL
    )

    if (is.null(fit_delta)) {
      delta_fail <- delta_fail + 1L
      results[i, c("delta_rl", "delta_lower", "delta_upper")] <- NA
    } else {
      ci <- delta_rl_CI(T, fit_delta)
      results[i, "delta_rl"]    <- ci$ReturnLevel
      results[i, "delta_lower"] <- ci$Lower
      results[i, "delta_upper"] <- ci$Upper
    }

    # Profile likelihood
    fit_prof <- tryCatch(
      suppressWarnings(gev_mle(bm_dummy, 1)),
      error = function(e) NULL
    )

    if (is.null(fit_prof) || any(is.na(fit_prof))) {
      profile_fail <- profile_fail + 1L
      results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
    } else {
      prof_return <- gev_return(fit_prof, m = T)
      prof_conf   <- tryCatch(
        suppressWarnings(confint(prof_return, profile = TRUE)),
        error = function(e) NULL
      )

      if (is.null(prof_conf)) {
        profileci_fail <- profileci_fail + 1L
        results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
      } else {
        results[i, "prof_rl"]    <- as.numeric(prof_return)
        results[i, "prof_lower"] <- prof_conf[1, 1]
        results[i, "prof_upper"] <- prof_conf[1, 2]
      }
    }
  }

  # Summary
  summary_df <- data.frame(
    ReturnPeriod = T,
    true_rl      = true_rl,

    delta_rl   = mean(results$delta_rl,  na.rm = TRUE),
    prof_rl    = mean(results$prof_rl,   na.rm = TRUE),

    RMSE_delta = sqrt(mean((results$delta_rl - true_rl)^2, na.rm = TRUE)),
    RMSE_prof  = sqrt(mean((results$prof_rl  - true_rl)^2, na.rm = TRUE)),

    Bias_delta     = mean(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof      = mean(results$prof_rl  - true_rl, na.rm = TRUE),

    Bias_delta_med = median(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof_med  = median(results$prof_rl  - true_rl, na.rm = TRUE),

    Coverage_delta = mean(results$delta_lower <= true_rl &
                          results$delta_upper >= true_rl, na.rm = TRUE),
    Coverage_prof  = mean(results$prof_lower  <= true_rl &
                          results$prof_upper  >= true_rl, na.rm = TRUE),
    
    delta_width    = mean(results$delta_upper - results$delta_lower, na.rm=TRUE),
    prof_width     = mean(results$prof_upper - results$prof_lower, na.rm=TRUE)
    
  )

  list(
    summary            = summary_df,
    raw                = results,        
    true_rl            = true_rl,
    delta_failures     = delta_fail,
    profile_failures   = profile_fail,
    profileci_failures = profileci_fail,
    delta_failure_rate   = delta_fail     / n_samples,
    profile_failure_rate = profile_fail   / n_samples
  )
}


coverage_plots <- function(raw, true_rl) {
  
  raw     <- raw
  true_rl <- true_rl
  
  n <- nrow(raw)
  raw$iteration <- seq_len(n)
  
  raw$delta_covers <- raw$delta_lower <= true_rl & raw$delta_upper >= true_rl
  raw$prof_covers  <- raw$prof_lower  <= true_rl & raw$prof_upper  >= true_rl
  
  # ── Delta plot ────────────────────────────────────────────────────────────────
  p_delta <- ggplot(raw, aes(x = iteration, colour = delta_covers)) +
  
    geom_linerange(aes(ymin = delta_lower, ymax = delta_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = delta_rl), shape = 16, size = 1.5, alpha = 0.5) +
  
    scale_colour_manual(
      name   = "Covers true Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", hjust = 0.5),
      legend.position  = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin    = margin(4, 6, 4, 6)
    )
  
  # ── Profile plot ──────────────────────────────────────────────────────────────
  p_prof <- ggplot(raw, aes(x = iteration, colour = prof_covers)) +
  
    geom_linerange(aes(ymin = prof_lower, ymax = prof_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = prof_rl), shape = 16, size = 1.5) +
  
    scale_colour_manual(
      name   = "Covers true Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor  = element_blank(),
      plot.title        = element_text(face = "bold", hjust = 0.5),
      legend.position   = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin     = margin(4, 6, 4, 6)
    )
  
  list(
    delta   = p_delta,
    profile = p_prof
  )
}

```

# NS case
```{r No seasonality Weibull}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0, 12)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
nseg_weibull <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
noseasonality25_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
noseasonality50_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
noseasonality100_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(noseasonality25_weibull, 
     noseasonality50_weibull, 
     noseasonality100_weibull,  
     nseg_weibull,
     file = "no_seasonality_weibul.RData")

```

```{r NO seasonality Data}
load("no_seasonality_weibul.RData")

noseasonality_weibull <- rbind(
  noseasonality25_weibull$summary,
  noseasonality50_weibull$summary,
  noseasonality100_weibull$summary
)

noseasonality_weibull
plot(nseg_weibull[1:(5*365)])
nsbm_weibull <- block_max(nseg_weibull, 365)
nsfit_weibull <- gev.fit(nsbm_weibull)
gev.diag(nsfit_weibull)

noseasonality100_weibull

plots <- coverage_plots(
  noseasonality100_weibull$raw, 
  noseasonality100_weibull$summary$true_rl
  )

plots$delta
plots$profile
```

# S1 
```{r v1 seasonality Weibull}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
V1seg_weibull <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
V1seasonality25_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
V1seasonality50_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
V1seasonality100_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(V1seasonality25_weibull, 
     V1seasonality50_weibull, 
     V1seasonality100_weibull,  
     V1seg_weibull,
     file = "V1_seasonality_weibul.RData")

```

```{r V1 seasonal Data}
load("V1_seasonality_weibul.RData")

V1seasonality_weibull <- rbind(
  V1seasonality25_weibull$summary,
  V1seasonality50_weibull$summary,
  V1seasonality100_weibull$summary
)

V1seasonality_weibull
plot(V1seg_weibull[1:(5*365)])
V1sbm_weibull <- block_max(V1seg_weibull, 365)
V1sfit_weibull <- gev.fit(V1sbm_weibull)
gev.diag(V1sfit_weibull)

V1seasonality100_weibull

plots <- coverage_plots(
  V1seasonality100_weibull$raw, 
  V1seasonality100_weibull$summary$true_rl
  )

plots$delta
plots$profile
```

# S2
```{r v2 seasonality Weibull}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,4,8,13,20,20,13,8,4,2,1)
sigma       <- rep(1, 12)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
V2seg_weibull <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
V2seasonality25_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
V2seasonality50_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
V2seasonality100_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(V2seasonality25_weibull, 
     V2seasonality50_weibull, 
     V2seasonality100_weibull,  
     V2seg_weibull,
     file = "V2_seasonality_weibul.RData")

```

```{r V2 Seasonal Data}
load("V2_seasonality_weibul.RData")

V2seasonality_weibull <- rbind(
  V2seasonality25_weibull$summary,
  V2seasonality50_weibull$summary,
  V2seasonality100_weibull$summary
)

V2seasonality_weibull
plot(V2seg_weibull[1:(5*365)])
V2sbm_weibull <- block_max(V2seg_weibull, 365)
V2sfit_weibull <- gev.fit(V2sbm_weibull)
gev.diag(V2sfit_weibull)

V2seasonality100_weibull

plots <- coverage_plots(
  V2seasonality100_weibull$raw, 
  V2seasonality100_weibull$summary$true_rl
  )

plots$delta
plots$profile
```

# S3 THIS IS THE CORRECT WAY AROUND 
Due to having run this at a later date
```{r v3 seasonality Weibull}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- rep(0,12)
sigma       <- c(1,2,3,4,5,6,6,5,4,3,2,1)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
V3seg_weibull <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
V3seasonality25_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
V3seasonality50_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
V3seasonality100_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(V3seasonality25_weibull, 
     V3seasonality50_weibull, 
     V3seasonality100_weibull,  
     V3seg_weibull,
     file = "V3_seasonality_weibul.RData")

```

```{r V3 seasonal Data}
load("V3_seasonality_weibul.RData")

V3seasonality_weibull <- rbind(
  V3seasonality25_weibull$summary,
  V3seasonality50_weibull$summary,
  V3seasonality100_weibull$summary
)

V3seasonality_weibull
plot(V3seg_weibull[1:(5*365)])
V3sbm_weibull <- block_max(V3seg_weibull, 365)
V3sfit_weibull <- gev.fit(V3sbm_weibull)
gev.diag(V3sfit_weibull)

V3seasonality100_weibull

plots <- coverage_plots(
  V3seasonality100_weibull$raw, 
  V3seasonality100_weibull$summary$true_rl
  )

plots$delta
plots$profile
```

# S4 
```{r v4 seasonality Weibull}
# Code used to save into no seasonality case for the gumbel no seasonality case
################################################################################
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma       <- c(1,2,3,4,5,6,6,5,4,3,2,1)

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
V4seg_weibull <- x_samples(GEV_samples, mu, sigma)

set.seed(123)
V4seasonality25_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
V4seasonality50_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
V4seasonality100_weibull <- delta_prof(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(V4seasonality25_weibull, 
     V4seasonality50_weibull, 
     V4seasonality100_weibull,  
     V4seg_weibull,
     file = "V4_seasonality_weibul.RData")

```

```{r V4 seasonal Data}
load("V4_seasonality_weibul.RData")

V4seasonality_weibull <- rbind(
  V4seasonality25_weibull$summary,
  V4seasonality50_weibull$summary,
  V4seasonality100_weibull$summary
)

V4seasonality_weibull
plot(V4seg_weibull[1:(5*365)])
V4sbm_weibull <- block_max(V4seg_weibull, 365)
V4sfit_weibull <- gev.fit(V4sbm_weibull)
gev.diag(V4sfit_weibull)

V4seasonality100_weibull

plots <- coverage_plots(
  V4seasonality100_weibull$raw, 
  V4seasonality100_weibull$summary$true_rl
  )

plots$delta
plots$profile
```

# THIS IS WRONG WRONG WRONG BIG FAT TABLE
```{r THE WRONG Fat table}

load("no_seasonality_weibul.RData")

noseasonality_weibull <- rbind(
  noseasonality25_weibull$summary,
  noseasonality50_weibull$summary,
  noseasonality100_weibull$summary
)

load("V1_seasonality_weibul.RData")

V1seasonality_weibull <- rbind(
  V1seasonality25_weibull$summary,
  V1seasonality50_weibull$summary,
  V1seasonality100_weibull$summary
)

load("V2_seasonality_weibul.RData")

V2seasonality_weibull <- rbind(
  V2seasonality25_weibull$summary,
  V2seasonality50_weibull$summary,
  V2seasonality100_weibull$summary
)

load("V3_seasonality_weibul.RData")

V3seasonality_weibull <- rbind(
  V3seasonality25_weibull$summary,
  V3seasonality50_weibull$summary,
  V3seasonality100_weibull$summary
)

load("V4_seasonality_weibul.RData")

V4seasonality_weibull <- rbind(
  V4seasonality25_weibull$summary,
  V4seasonality50_weibull$summary,
  V4seasonality100_weibull$summary
)

prepare_table_delta <- function(x, scenario, method){
  
  if(method == "Delta"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        delta_rl,
        RMSE_delta,
        Bias_delta,
        Bias_delta_med,
        Coverage_delta,
        delta_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = delta_rl,
        RMSE = RMSE_delta,
        Bias = Bias_delta,
        `MD Bias` = Bias_delta_med,
        Coverage = Coverage_delta,
        Width = delta_width
      )
    
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}



prepare_table_prof <- function(x, scenario, method){
  
  if(method == "Profile"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        prof_rl,
        RMSE_prof,
        Bias_prof,
        Bias_prof_med,
        Coverage_prof,
        prof_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = prof_rl,
        RMSE = RMSE_prof,
        Bias = Bias_prof,
        `MD Bias` = Bias_prof_med,
        Coverage = Coverage_prof,
        Width = prof_width
      )
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}


delta_all <- rbind(
  prepare_table_delta(noseasonality_weibull, "NS", "Delta"),
  prepare_table_delta(V1seasonality_weibull, "S1", "Delta"),
  prepare_table_delta(V2seasonality_weibull, "S2", "Delta"),
  prepare_table_delta(V4seasonality_weibull, "S3", "Delta"),
  prepare_table_delta(V3seasonality_weibull, "S4", "Delta")
)

prof_all <- rbind(
  prepare_table_prof(noseasonality_weibull, "NS", "Profile"),
  prepare_table_prof(V1seasonality_weibull, "S1", "Profile"),
  prepare_table_prof(V2seasonality_weibull, "S2", "Profile"),
  prepare_table_prof(V4seasonality_weibull, "S3", "Profile"),
  prepare_table_prof(V3seasonality_weibull, "S4", "Profile")
)

delta_all <- delta_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

prof_all <- prof_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )


delta_all
prof_all


delta_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Delta method for Return level estimation performance for weibull.",
    label = "tab:estimation"
  ) |>
  save_kable("all_seasonality_weibull_delta.tex")

prof_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile Likelihood method for Return level estimation performance for weibull.",
    label = "tab:estimation"
  ) |>
  save_kable("all_seasonality_weibull_prof.tex")





```

# THE RIGHT BIG FAT TABLE (Standardised)
```{r Standardised version}

load("no_seasonality_weibul.RData")

noseasonality_weibull <- rbind(
  noseasonality25_weibull$summary,
  noseasonality50_weibull$summary,
  noseasonality100_weibull$summary
)

load("V1_seasonality_weibul.RData")

V1seasonality_weibull <- rbind(
  V1seasonality25_weibull$summary,
  V1seasonality50_weibull$summary,
  V1seasonality100_weibull$summary
)

load("V2_seasonality_weibul.RData")

V2seasonality_weibull <- rbind(
  V2seasonality25_weibull$summary,
  V2seasonality50_weibull$summary,
  V2seasonality100_weibull$summary
)

load("V3_seasonality_weibul.RData")

V3seasonality_weibull <- rbind(
  V3seasonality25_weibull$summary,
  V3seasonality50_weibull$summary,
  V3seasonality100_weibull$summary
)

load("V4_seasonality_weibul.RData")

V4seasonality_weibull <- rbind(
  V4seasonality25_weibull$summary,
  V4seasonality50_weibull$summary,
  V4seasonality100_weibull$summary
)

########################################

LS_weibull <- tibble::tribble(
  ~Scenario, ~L, ~S,
  "NS",       0,  1,
  "S1",       6,  1,
  "S2",      20,  1,
  "S3",       0,  6,
  "S4",       6,  6
)

## ---------------------------------------------------------------
## 2. Pull delta and profile results into one row per scenario/T
## ---------------------------------------------------------------

prepare_combined <- function(x, scenario) {
  x |>
    select(
      ReturnPeriod,
      true_rl,
      prof_rl, delta_rl,
      RMSE_prof, RMSE_delta,
      Bias_prof, Bias_prof_med,
      Coverage_delta, delta_width,
      Coverage_prof,  prof_width
    ) |>
    mutate(Scenario = scenario, .before = 1)
}

combined_raw <- rbind(
  prepare_combined(noseasonality_weibull, "NS"),
  prepare_combined(V1seasonality_weibull, "S1"),
  prepare_combined(V2seasonality_weibull, "S2"),
  prepare_combined(V3seasonality_weibull, "S3"),   # the right way
  prepare_combined(V4seasonality_weibull, "S4")    # the right way
)

## Sanity check: point estimates should agree between methods
max_diff <- max(abs(combined_raw$prof_rl   - combined_raw$delta_rl),
                abs(combined_raw$RMSE_prof - combined_raw$RMSE_delta))
if (max_diff > 5e-4) {
  warning("Delta and profile point estimates differ by up to ",
          round(max_diff, 4),
          ". Table reports the profile values only -- check before submitting.")
}

## ---------------------------------------------------------------
## 3. Build the table, optionally standardised
## ---------------------------------------------------------------
build_weibull_table <- function(dat, LS, standardise = TRUE) {

  dat |>
    left_join(LS, by = "Scenario") |>
    mutate(
      shift = if (standardise) L else 0,
      scale = if (standardise) S else 1
    ) |>
    transmute(
      Scenario,
      T        = ReturnPeriod,
      Ref      = (true_rl - shift) / scale,
      Est      = (prof_rl - shift) / scale,
      RMSE     = RMSE_prof     / scale,
      Bias     = Bias_prof     / scale,
      `M Bias` = Bias_prof_med / scale,
      DCovers  = Coverage_delta,
      DWidth   = delta_width   / scale,
      PCovers  = Coverage_prof,
      PWidth   = prof_width    / scale
    )
}

print_weibull_table <- function(tab, caption, label, file,
                               digits = 3,
                               digits_by_col = NULL) {

  ## default: same precision everywhere
  num_cols <- names(tab)[vapply(tab, is.numeric, logical(1))]
  num_cols <- setdiff(num_cols, "T")

  dg <- setNames(rep(digits, length(num_cols)), num_cols)
  if (!is.null(digits_by_col)) {
    dg[names(digits_by_col)] <- digits_by_col
  }

  chr <- tab
  for (nm in num_cols) {
    chr[[nm]] <- sprintf(paste0("%.", dg[[nm]], "f"), chr[[nm]])
  }
  chr$T <- as.character(chr$T)

  chr$Scenario[duplicated(chr$Scenario)] <- ""

  grp <- "& & \\multicolumn{2}{c}{Return Level} & \\multicolumn{3}{c}{Error} & \\multicolumn{2}{c}{Delta} & \\multicolumn{2}{c}{Profile} \\\\"
  mid <- "\\cmidrule(lr){3-4} \\cmidrule(lr){5-7} \\cmidrule(lr){8-9} \\cmidrule(lr){10-11}"
  nms <- "Scenario & T & Ref & Est & RMSE & Bias & M Bias & Covers & Width & Covers & Width \\\\"

  header <- paste("\\toprule", grp, mid, nms, "\\midrule", sep = "\n")

  n_per  <- 3L
  breaks <- seq(n_per, nrow(chr) - n_per, by = n_per)

  addtorow <- list(
    pos     = c(list(-1), as.list(breaks), list(nrow(chr))),
    command = c(header, rep("\\midrule\n", length(breaks)), "\\bottomrule\n")
  )

  xt <- xtable(chr,
               caption = caption,
               label   = label,
               align   = paste0("ll", strrep("r", ncol(chr) - 1)))

  print(xt,
        file              = file,
        include.rownames  = FALSE,
        include.colnames  = FALSE,
        hline.after       = NULL,
        add.to.row        = addtorow,
        booktabs          = TRUE,
        caption.placement = "top",
        sanitize.text.function = identity,
        table.placement   = "ht")
}

std_table <- build_weibull_table(combined_raw, LS_weibull, standardise = TRUE)

print_weibull_table(
  std_table,
  caption = "...",
  label   = "tab:weibull_standardised",
  file    = "weibull_standardised.tex",
  digits  = 3,
  digits_by_col = c(Ref = 3, Est = 3, DCovers = 3, PCovers = 3)
)

std_table

```

# This is the PREVIOUS VERSION OF noseasonality case for weibull
NOTE WE SHOULD CHANGE THE INSIDE CODE FOR x_samples AS REQUIRED  
Tested for sets of 100 year samples instead due to failure of optimisation 
This is also a version with old sigma that was not of a comparable version to the Gumbel case
```{r no seasonality case weibull, eval=FALSE}
# Code used to save into no seasonality case for the weibull no seasonality case
################################################################################
set.seed(123)

GEV_samples <- 100 * 365
n_samples <- 100 # number of n_samples 
q_samples <- 100000 * 365

rlcolnames <- c("rl10", "rl20", "rl100", "rl1000")
T <- c(10,20,100,1000)  

mu <- rep(1,12)
sigma <- rep(10,12)

nseg_weibull <- x_samples(GEV_samples, mu, sigma)
nsbm_weibull <- block_max(nseg_weibull, 365)
nsfit_weibull <- gev.fit(nsbm_weibull)
nsdiag_weibull <- gev.diag(nsfit_weibull)

noseasonality_weibull <- delta_prof(GEV_samples, n_samples, q_samples, rlcolnames, T, mu, sigma)
save(noseasonality_weibull, nseg_weibull, nsbm_weibull, nsfit_weibull, nsdiag_weibull, file = "no_seasonality_weibull.RData")
################################################################################
# To print the table from what we have saved 

load("no_seasonality_weibull.RData")
noseasonality_weibull
plot(nseg_weibull[1:(5*365)]) 
```

# 100 year version of S2 with the old sigma
```{r seasonal case, eval=FALSE}
# Code used to save into no seasonality case for the weibull no seasonality case
################################################################################
set.seed(123)

GEV_samples <- 100 * 365 # Used 100 here and thats better 
n_samples <- 100 # number of n_samples 
q_samples <- 100000 * 365

rlcolnames <- c("rl10", "rl20", "rl100", "rl1000")
T <- c(10,20,100,1000)  

mu <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma <- rep(10,12)

seg_weibull <- x_samples(GEV_samples, mu, sigma)
sbm_weibull <- block_max(seg_weibull, 365)
sfit_weibull <- gev.fit(sbm_weibull)
sdiag_weibull <- gev.diag(sfit_weibull)

set.seed(123)

seasonal_weibull <- delta_prof(GEV_samples, n_samples, q_samples, rlcolnames, T, mu, sigma)

save(seasonal_weibull, seg_weibull, sbm_weibull, sfit_weibull, sdiag_weibull, file = "seasonal_weibull.RData")
################################################################################
# To print the table from what we have saved 

load("seasonal_weibull.RData")
seasonal_weibull
plot(seg_weibull[1:(5*365)])
```

# Another version
```{r seasonal case, eval=FALSE}
# Code used to save into no seasonality case for the weibull no seasonality case
################################################################################
set.seed(123)

GEV_samples <- 100 * 365
n_samples <- 100 # number of n_samples 
q_samples <- 100000 * 365

rlcolnames <- c("rl10", "rl20", "rl100", "rl1000")
T <- c(10,20,100,1000)  

mu <- c(5,6,8,12,16,20,25,25,20,15,10,6)
sigma <- rep(10,12)

s2eg_weibull <- x_samples(GEV_samples, mu, sigma)
s2bm_weibull <- block_max(s2eg_weibull, 365)
s2fit_weibull <- gev.fit(s2bm_weibull)
s2diag_weibull <- gev.diag(s2fit_weibull)

set.seed(123)

seasonal2_weibull <- delta_prof(GEV_samples, n_samples, q_samples, rlcolnames, T, mu, sigma)

save(seasonal2_weibull, s2eg_weibull, s2bm_weibull, s2fit_weibull, s2diag_weibull, file = "seasonal2_weibull.RData")
################################################################################
# To print the table from what we have saved 

load("seasonal2_weibull.RData")
s2fit_weibull
seasonal2_weibull
plot(s2eg_weibull[1:(5*365)])
plot(s2eg_weibull)
plot(s2bm_weibull)
```
