---
title: "M4A Combination"
author: ''
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ismev) # Coles Package
library(evmissing) # Simpson & Northrop package 
library(readr) # for csv and things 
library(ggplot2)
library(dplyr)
library(ggrepel)
library(patchwork)
library(tidyr)
```

# Generally for generating Figures
This markdown document is generated generally for the purpose of creating figures for MSc dissertation in Statistics in extremes 

## Generating Red Sea seasonal observations structure for a singular year
```{r Generating Red Sea seasonal observations structure}

# Requires gumbel markdown initialisation stuff or others depending on the domain of attraction you want

month_days  <- c(31,28,31,30,31,30,31,31,30,31,30,31)
month_index <- rep(1:12, times = month_days)

scenarios <- list(
  S1 = list(mu = c(1,2,3,4,5,6,6,5,4,3,2,1),      sigma = rep(1, 12)),
  S2 = list(mu = c(1,2,4,8,13,20,20,13,8,4,2,1),  sigma = rep(1, 12)),
  S3 = list(mu = rep(0, 12),                      sigma = c(1,2,3,4,5,6,6,5,4,3,2,1)),
  S4 = list(mu = c(1,2,3,4,5,6,6,5,4,3,2,1),      sigma = c(1,2,3,4,5,6,6,5,4,3,2,1))
)

set.seed(123)
block_df <- bind_rows(lapply(names(scenarios), function(s) {
  p <- scenarios[[s]]
  data.frame(
    scenario = s,
    day      = 1:365,
    x        = x_samples(365, p$mu, p$sigma),
    mu       = p$mu[month_index],
    sigma    = p$sigma[month_index]
  )
}))
block_df$scenario <- factor(block_df$scenario, levels = names(scenarios))

# the block maximum — the quantity the GEV actually sees
bm_df <- block_df %>% group_by(scenario) %>% slice_max(x, n = 1, with_ties = FALSE)

normsblocks <- ggplot(block_df, aes(x = day)) +
  # geom_ribbon(aes(ymin = mu - 2 * sigma, ymax = mu + 2 * sigma), fill = "grey", alpha = 0.3) +
  geom_point(aes(y = x), size = 2, alpha = 1, colour = "lightblue") +
  geom_point(data = bm_df, aes(y = x), colour = "black", size = 2.2) +
  # geom_text(data = bm_df, aes(y = x, label = "block max"), colour = "black", vjust = -0.8, size = 3) +
  facet_wrap(~ scenario, ncol = 2) +
  scale_x_continuous(
    name   = "Day of year",
    breaks = cumsum(month_days) - month_days / 2,
    labels = c("J","F","M","A","M","J","J","A","S","O","N","D"),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(name = "Daily observation") +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text       = element_text(),
  )

normsblocks


```

## Generating Red Sea seasonal observations structure for 3 years 
```{r Generating Red Sea seasonal observations structure for 3 years}

month_days  <- c(31,28,31,30,31,30,31,31,30,31,30,31)
month_index <- rep(1:12, times = month_days)

n_years <- 3
n_days  <- 365 * n_years

scenarios <- list(
  S1 = list(mu = c(1,2,3,4,5,6,6,5,4,3,2,1),      sigma = rep(1, 12)),
  S2 = list(mu = c(1,2,4,8,13,20,20,13,8,4,2,1),  sigma = rep(1, 12)),
  S3 = list(mu = rep(0, 12),                      sigma = c(1,2,3,4,5,6,6,5,4,3,2,1)),
  S4 = list(mu = c(1,2,3,4,5,6,6,5,4,3,2,1),      sigma = c(1,2,3,4,5,6,6,5,4,3,2,1))
)

set.seed(123)
block_df <- bind_rows(lapply(names(scenarios), function(s) {
  p <- scenarios[[s]]
  data.frame(
    scenario = s,
    day      = 1:n_days,
    year     = rep(seq_len(n_years), each = 365),
    x        = x_samples(n_days, p$mu, p$sigma),
    mu       = rep(p$mu[month_index],    n_years),
    sigma    = rep(p$sigma[month_index], n_years)
  )
}))
block_df$scenario <- factor(block_df$scenario, levels = names(scenarios))

# one block maximum per year per scenario
bm_df <- block_df %>%
  group_by(scenario, year) %>%
  slice_max(x, n = 1, with_ties = FALSE) %>%
  ungroup()

# month-centre positions repeated across the three years
month_centres <- cumsum(month_days) - month_days / 2
x_breaks <- as.vector(outer(month_centres, 365 * (0:(n_years - 1)), "+"))
x_labels <- rep(c("J","F","M","A","M","J","J","A","S","O","N","D"), n_years)

ggplot(block_df, aes(x = day)) +
  geom_vline(xintercept = 365 * seq_len(n_years - 1) + 0.5,
             linetype = "dashed", colour = "grey60", linewidth = 0.4) +
  geom_point(aes(y = x), size = 1.1, alpha = 1, colour = "lightblue") +
  geom_point(data = bm_df, aes(y = x), colour = "black", size = 2.2) +
  facet_wrap(~ scenario, ncol = 1) +
  scale_x_continuous(
    name   = "Day of year",
    breaks = x_breaks,
    labels = x_labels,
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(name = "Daily observation") +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text       = element_text(),
    axis.text.x      = element_text(size = 7)
  )

```

## Data cleaning for 108 locations 
```{r Producing big graph for the 108 locations}

# this is location 1 
load("Test_run50_Red_Sea.RData")

test50_redsea <- rbind(
  test50_redsea_25$summary, 
  test50_redsea_50$summary,
  test50_redsea_100$summary
)


for (i in 2:5) {
  
  load(paste0("loc", i, "_Red_Sea.RData"))
  
  summary_df <- rbind(
    get(paste0("loc", i, "_redsea_25"))$summary,
    get(paste0("loc", i, "_redsea_50"))$summary,
    get(paste0("loc", i, "_redsea_100"))$summary
  )

  assign(paste0("loc", i, "_redsea"), summary_df)
  
  print(summary_df)
}

for (i in 6:52) {

  load(paste0("location", i, "_Red_Sea.RData"))

  summary_df <- rbind(
    get(paste0("loc", i, "_redsea_25"))$summary,
    get(paste0("loc", i, "_redsea_50"))$summary,
    get(paste0("loc", i, "_redsea_100"))$summary
  )

  assign(paste0("loc", i, "_redsea"), summary_df)
  
  print(summary_df)
}

for (i in 53:108) {

  load(paste0("location", i, "_Red_Sea.RData"))

  summary_df <- rbind(
    get(paste0("loc", i, "_yredsea_25"))$summary,
    get(paste0("loc", i, "_yredsea_50"))$summary,
    get(paste0("loc", i, "_yredsea_100"))$summary
  )

  assign(paste0("loc", i, "_redsea"), summary_df)
  
  print(summary_df)
}


redsea_all <- list()

# Location 1
redsea_all[[1]] <- test50_redsea %>%
  mutate(Location = 1)

# Locations 2-108
for(i in 2:108){

  df <- get(paste0("loc", i, "_redsea"))

  redsea_all[[i]] <- df %>%
    mutate(Location = i)

}

redsea_all <- bind_rows(redsea_all)

red25 <- filter(redsea_all, ReturnPeriod == 25)
red50 <- filter(redsea_all, ReturnPeriod == 50)
red100 <- filter(redsea_all, ReturnPeriod == 100)

```

## Generating the plots for the 108 locations 
```{r Plots for 108 locations}

p1 <- ggplot(red100,
       aes(y = Coverage_prof,
           x = Location)) +

  geom_point(size = 1.8, colour = "steelblue") +

  geom_hline(yintercept = 0.95,
             linetype = 2,
             colour = "black") +

  labs(
    y = "Coverage",
    x = "Location"
  ) +

  theme_bw()

ylim <- range(
    c(red100$Bias_prof,
      red100$Bias_prof_med)
)


p2 <- ggplot(red100,
       aes(y=Bias_prof,
           x=Location)) +

  geom_point(colour = "steelblue") +

  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "black") +
  
  scale_y_continuous(limits = ylim) +
  
  labs(
    y = "Mean Bias",
    x = "Location"
  ) +

  theme_bw()

bias_df <- pivot_longer(
  red100,
  cols = c(Bias_prof, Bias_prof_med),
  names_to = "Statistic",
  values_to = "Bias"
)

bias_df$Statistic <- factor(
  bias_df$Statistic,
  levels = c("Bias_prof", "Bias_prof_med"),
  labels = c("Mean Bias", "Median Bias")
)

bias_combo <- ggplot(bias_df,
       aes(Location, Bias)) +

  geom_point(colour = "steelblue") +

  geom_hline(yintercept = 0,
             linetype = 2) +

  facet_wrap(~Statistic, ncol = 2) +

  theme_bw()+
  theme(
  strip.background = element_blank(),
  strip.text = element_text()
)

p3 <- ggplot(red100,
       aes(y=Bias_prof_med,
           x=Location)) +

  geom_point(colour = "steelblue") +

  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "black") +
  
  scale_y_continuous(limits = ylim) +
  
  labs(
    y = "Median Bias",
    x = "Location"
  ) +

  theme_bw() 

p4 <- ggplot(red100,
       aes(y = prof_width,
           x = Location)) +

  geom_point(colour = "steelblue") +
  
  labs(
    y = "Interval Width",
    x = "Location"
  ) +

  theme_bw()

p5 <- ggplot(red100,
       aes(y=true_rl,
           x=Location)) +

  geom_point(colour = "steelblue", size = 2) +
  
  labs(
    y = "Reference Return Level",
    x = "Location"
  ) +

  theme_bw()

p1 / p4
p5
bias_combo

#############################################################################
# Produce the map of the coverage and the interval width 

coords <- as.data.frame(loc) |>
  setNames(c("lon", "lat")) |>
  mutate(Location = row_number())

map_df <- left_join(red100, coords, by = "Location")
stopifnot(!any(is.na(map_df$lon)))

lim_cov <- 0.95 + c(-1, 1) * max(abs(map_df$Coverage_prof - 0.95))

base_map <- list(
  coord_quickmap(),
  theme_bw(),
  theme(panel.grid = element_blank(),
        legend.position = "right",
        legend.key.height = unit(1.1, "cm"))
)

p_cov <- ggplot(map_df, aes(lon, lat, fill = Coverage_prof)) +
  geom_point(shape = 21, size = 4.5, stroke = 0.25, colour = "grey25") +
  scale_fill_gradient2(
    name = "Coverage", midpoint = 0.95, limits = lim_cov,
    low = "steelblue", mid = "grey95", high = "red"
  ) +
  labs(x = "Longitude", y = "Latitude") +
  base_map +
  theme_bw()

p_wid <- ggplot(map_df, aes(lon, lat, fill = prof_width)) +
  geom_point(shape = 21, size = 4.5, stroke = 0.25, colour = "grey25") +
  scale_fill_viridis_c(name = "Interval\nwidth", option = "mako", direction = -1) +
  labs(x = "Longitude", y = NULL) +
  base_map +
  theme_bw()

# two by one graph with scale filled for coverage and interval width 
p_cov | p_wid

```

## Triple threat of the 3 methodologies visualisation of 5 year/block samples
```{r block maxima, POT, r largest}

set.seed(123)
x <- rnorm(5*365, 0, 1)

example <- data.frame(Index = 1:(5*365), Observations = x)
example$Block <- (example$Index - 1) %/% 365 + 1

block_lines <- seq(365, 365*4, by = 365)
x_scale <- scale_x_continuous(breaks = seq(365/2, 365*5, by = 365),
                              labels = paste(1:5))

base <- function(p) {
  p +
    geom_vline(xintercept = block_lines, alpha = 0.5, linetype = "dashed") +
    scale_y_continuous(expand = expansion(mult = c(0.00, 0.12))) +
    x_scale +
    labs(x = "Year Index", y = "Data") +
    theme_bw(base_size = 14)
}

# ---- Block maxima ----------------------------------------------------------
max_points <- example |>
  group_by(Block) |>
  slice_max(Observations, n = 1) |>
  ungroup()

p_bm <- base(
  ggplot(example, aes(Index, Observations)) +
    geom_point(size = 1.5, colour = "lightblue") +
    geom_point(data = max_points, colour = "black", size = 3)
) +
  geom_text_repel(data = max_points,
                  aes(label = paste("Block", Block, "max")),
                  colour = "black", fontface = "bold",
                  box.padding = 0.4, point.padding = 0.2, seed = 123) +
  ggtitle("Block maxima")

# ---- r largest order statistics --------------------------------------------
r <- 5

r_points <- example |>
  group_by(Block) |>
  slice_max(Observations, n = r) |>
  ungroup()

p_r <- base(
  ggplot(example, aes(Index, Observations)) +
    geom_point(size = 1.5, colour = "lightblue") +
    geom_point(data = r_points, colour = "black", size = 2.2)
) +
  ggtitle(paste0("r largest order statistics (r = ", r, ")"))

# ---- Peaks over threshold --------------------------------------------------
# u <- quantile(x, 0.95)
u <- 2

pot_points <- filter(example, Observations > u)

p_pot <- base(
  ggplot(example, aes(Index, Observations)) +
    geom_point(size = 1.5, colour = "lightblue") +
    geom_point(data = pot_points, colour = "black", size = 2.2)
) +
  geom_hline(yintercept = u, colour = "black", linewidth = 1) +
  # annotate("text", x = 5, y = u, label = "u", hjust = 0, vjust = -0.5, fontface = "bold", size = 5) +
  ggtitle("Peaks over threshold")

p_bm <- p_bm + labs(x = NULL) 
p_r  <- p_r  + labs(x = NULL)

p_bm / p_r / p_pot
```

## Block maxima visualisation with the GEV fitted on top of the histogram
```{r Dummy case for producing a samples from random normal to use in block maxima}
set.seed(123)

x <- rnorm(5*365, 0, 1)

block_max <- function(df, block_size){

  df$Block <- (df$Index - 1) %/% block_size + 1

  maxima <- df |>
    dplyr::group_by(Block) |>
    dplyr::slice_max(Observations, n = 1) |>
    dplyr::ungroup()

  maxima
}

example <- data.frame(
  Index = 1:(5*365),
  Observations = x
)

max_points <- block_max(example, 365)

bm_example <- ggplot(example, aes(Index, Observations)) +

  geom_point(size = 1.5, colour = "lightblue") +

  geom_vline(
    xintercept = seq(365, 365*4, by = 365),
    alpha = 0.5, 
    linetype = "dashed"
  ) +

  geom_point(
    data = max_points,
    colour = "black",
    size = 3
  ) +

  geom_text_repel(
    data = max_points,
    aes(label = paste("Block", Block, "max")),
    colour = "black",
    fontface = "bold",
    box.padding = 0.4,
    point.padding = 0.2,
    seed = 123
  ) +
  
  scale_y_continuous(
    expand = expansion(mult = c(0.00, 0.12))
  ) +
  
  scale_x_continuous(
    breaks = seq(365/2, 365 * 5, by = 365),
    labels = paste(1:5)
  ) +

  labs(
    x = "Year Index",
    y = "Data"
  ) +
  
  # Depends on what theme 
  theme_bw(base_size=14)

###############################################################################

set.seed(123)

x <- rnorm(50*365, 0, 1)

example50 <- data.frame(
  Index = 1:(50*365),
  Observations = x
)

max_points <- block_max(example50, 365)

bm50_example <- ggplot(example50, aes(Index, Observations)) +

  geom_point(
    aes(colour = "Observations"),
    size = 1.5) +

  geom_vline(
    xintercept = seq(365, 365*49, by = 365),
    alpha = 0.5, 
    linetype = "dashed"
  ) +

  geom_point(
    aes(colour = "Annual maxima"),
    data = max_points,
    size = 3
  ) +
  
  scale_colour_manual(
    values = c(
      "Observations" = "lightblue",
      "Annual maxima" = "black")
    ) +

  labs(
    x = "Observation Index",
    y = "Data"
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(
        size = c(3,3)
      )
    )
  ) +

  labs(
    x = "Observation Index",
    y = "Data",
    colour = NULL
  ) +
  
  theme_bw(base_size = 14) +

  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor  = element_blank(),
    plot.title        = element_text(face = "bold", hjust = 0.5),

    legend.position = c(0.75,0.10),
    legend.background = element_rect(
      fill = "white",
      colour = "grey70"
    ),
    legend.margin = margin(4,6,4,6)
  )

bm50_example

# 
################################################################################

example_max <- ggplot(max_points, aes(Index, Observations)) +
  geom_point(
    data = max_points,
    colour = "black",
    size = 3
  ) +
  
  labs(
    x = "Block Maxima",
    y = "Data"
  ) +
  
  
  theme_bw(base_size=14)


###############################################################################

fit <- gev.fit(max_points$Observations, show = FALSE)

mu    <- fit$mle[1]
sigma <- fit$mle[2]
xi    <- fit$mle[3]

dgev <- function(x, mu, sigma, xi){

  if(abs(xi) < 1e-8){

    z <- (x - mu)/sigma

    return(exp(-(z + exp(-z)))/sigma)

  }

  y <- 1 + xi*(x-mu)/sigma

  dens <- ifelse(
    y > 0,
    (1/sigma) *
      y^(-1/xi - 1) *
      exp(-y^(-1/xi)),
    0
  )

  dens

}

x_grid <- seq(min(max_points$Observations),
              max(max_points$Observations),
              length.out = 500)

gev_curve <- data.frame(
  x = x_grid,
  density = dgev(x_grid, mu, sigma, xi)
)

points_df <- data.frame(
  x = max_points$Observations,
  y = 0
)

example_hist <- ggplot() +

  geom_histogram(
    data = data.frame(max_points$Observations),
    aes(max_points$Observations, after_stat(density)),
    bins = 9,
    fill = "lightblue",
    colour = "black"
  ) +

  geom_line(
    data = gev_curve,
    aes(x, density),
    linewidth = 1.2,
    colour = "black"
  ) +

  geom_point(
    data = points_df,
    aes(x = x, y = y),
    size = 2
  ) +
  
  labs(
    x = "Annual maxima",
    y = "Density"
  ) +

  theme_bw(base_size = 14)

example_hist

##############################################################################

# bm_example | bm50_example

# example_max | example_hist

# bm_example

# bm50_example | example_hist


```

## For likelihood visualisation of skewed return level
```{r For likelihood visualisation}

fit    <- gev.fit(max_points$Observations)
fit_ev <- gev_mle(max_points$Observations, 1)

x <- confint(fit_ev, profile = TRUE)

gev.prof(fit, 100, 3.5, 5, conf = 0.95, nint = 100)


library(ggplot2)

z <- max_points$Observations
T <- 100

fit <- gev.fit(z, show = FALSE)
ci_delta <- delta_rl_CI(T, fit) 

# Set up grid for profile log-likelihood return levels
prof_ll <- function(zp, z, T) {
  y <- -log(1 - 1/T)
  nll <- function(par) {
    sig <- exp(par[1]); xi <- par[2]
    mu  <- if (abs(xi) < 1e-8) zp + sig*log(y) else zp + (sig/xi)*(1 - y^(-xi))
    w   <- 1 + xi*(z - mu)/sig
    if (sig <= 0 || any(w <= 0)) return(1e10)
    if (abs(xi) < 1e-8)
      length(z)*log(sig) + sum((z-mu)/sig) + sum(exp(-(z-mu)/sig))
    else
      length(z)*log(sig) + (1+1/xi)*sum(log(w)) + sum(w^(-1/xi))
  }
  -optim(c(log(sd(z)), 0.1), nll)$value
}

# Manually adjusted this until it looks good 
grid <- seq(3.4875, 4.5, length.out = 150) 
ll   <- vapply(grid, prof_ll, numeric(1), z = z, T = T)

max_ll <- max(ll)
cutoff <- max_ll - qchisq(0.95, 1)/2

inside <- grid[ll >= cutoff]
ci_prof <- range(inside)

prof_df <- data.frame(rl = grid, ll = ll)

prof_rl <- ggplot(prof_df, aes(rl, ll)) +
  geom_line(linewidth = 0.7) +
  geom_hline(yintercept = cutoff, colour = "grey40", linetype = "dashed") +

  annotate("segment", x = ci_prof[1], xend = ci_prof[2],
           y = cutoff, yend = cutoff,
           colour = "steelblue", linewidth = 1.4) +
  annotate("point", x = ci_prof, y = c(cutoff, cutoff),
           colour = "steelblue", size = 2.5) +

  annotate("segment", x = ci_delta$Lower, xend = ci_delta$Upper,
           y = cutoff - 0.35, yend = cutoff - 0.35,
           colour = "grey30", linewidth = 1.4) +
  annotate("point", x = c(ci_delta$Lower, ci_delta$Upper),
           y = rep(cutoff - 0.35, 2), colour = "grey30", size = 2.5) +

  annotate("text", x = ci_prof[2], y = cutoff - 0.15,
           label = "Profile likelihood", hjust = 1, colour = "steelblue", size = 5) +
  annotate("text", x = ci_delta$Upper, y = cutoff - 0.5,
           label = "Delta method", hjust = 1, colour = "steelblue", size = 5) +

  geom_vline(xintercept = ci_delta$ReturnLevel, linetype = "solid", alpha = 0.6) +
  labs(x = "100 year return level", y = "Profile log-likelihood") +
  theme_bw()

prof_rl

```

## Return level against the log y_p plot 
```{r For examples of the linear relationship between return level and log yp}

gev_rl <- function(p, mu, sigma, xi) {
  y <- -log(1 - p)
  if (abs(xi) < 1e-6) mu - sigma * log(y)
  else mu + (sigma / xi) * (y^(-xi) - 1)
}

# Initialising the values of interest
mu    <- 0
sigma <- 1
xis   <- c(-0.2, 0, 0.2)
probs <- seq(0.01, 0.999, length.out = 500)

plot_df <- do.call(rbind, lapply(xis, function(xi) {
  data.frame(
    log_yp = -log(-log(1 - probs)),
    rl     = sapply(probs, gev_rl, mu = mu, sigma = sigma, xi = xi),
    xi     = xi
  )
}))

plot_df$xi <- factor(plot_df$xi,
                     levels = c(-0.2, 0, 0.2),
                     labels = c("ξ = -0.2 (Weibull)",
                                "ξ = 0 (Gumbel)",
                                "ξ = 0.2 (Fréchet)"))

label_df <- do.call(rbind, lapply(xis, function(xi) {
  p <- 0.01
  data.frame(
    log_yp = -log(-log(1 - p)),
    rl     = gev_rl(p, mu, sigma, xi),
    xi     = xi
  )
}))

label_df$xi <- factor(label_df$xi,
                      levels = c(-0.2, 0, 0.2),
                      labels = c("ξ = -0.2",
                                 "ξ = 0",
                                 "ξ = 0.2"))

xi_plot <- ggplot(plot_df, aes(x = log_yp, y = rl, group = xi)) +
  geom_line(linewidth = 0.8, colour = "black") +
  geom_text(
    data  = label_df,
    aes(label = xi),
    hjust = 0,
    nudge_x = 0.05,
    size  = 3.5
  ) +
  scale_x_continuous(
  name   = expression(-log~y[p]),
  breaks = seq(-2, 5, by = 1),
  expand = expansion(mult = c(0.00, 0.12))
  ) +
  scale_y_continuous(name = "Return Level") +
  theme_bw() +
  theme(
    legend.position  = "none",
    panel.grid.minor = element_blank()
  )

xi_plot

```

## Initial attempt at making the coverage plot before using it individually in the other markdown files
```{r coverage plot function tester}

coverage_plots <- function(raw, true_rl) {
  
  raw     <- raw
  true_rl <- true_rl
  
  n <- nrow(raw)
  raw$iteration <- seq_len(n)
  
  raw$delta_covers <- raw$delta_lower <= true_rl & raw$delta_upper >= true_rl
  raw$prof_covers  <- raw$prof_lower  <= true_rl & raw$prof_upper  >= true_rl
  
  # Delta plot 
  p_delta <- ggplot(raw, aes(x = iteration, colour = delta_covers)) +
  
    geom_linerange(aes(ymin = delta_lower, ymax = delta_upper),
                   linewidth = 0.25, alpha = 0.3) +
    geom_point(aes(y = delta_rl), shape = 16, size = 2, alpha = 0.7) +
  
    scale_colour_manual(
      name   = "Covers true Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level",
                       limits = c(2.8,11),
                       breaks = seq(3, 12, by = 2),
                       expand = expansion(mult = c(0.02, 0.02))) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", hjust = 0.5),
      legend.position  = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin    = margin(4, 6, 4, 6)
    )
  
  # Profile plot 
  p_prof <- ggplot(raw, aes(x = iteration, colour = prof_covers)) +
  
    geom_linerange(aes(ymin = prof_lower, ymax = prof_upper),
                   linewidth = 0.25, alpha = 0.3) +
    geom_point(aes(y = prof_rl), shape = 16, size = 2, alpha = 0.7) +
  
    scale_colour_manual(
      name   = "Covers true Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level",
                       limits = c(2.8,11),
                       breaks = seq(3, 12, by = 2),
                       expand = expansion(mult = c(0.02, 0.02))) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor  = element_blank(),
      plot.title        = element_text(face = "bold", hjust = 0.5),
      legend.position   = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin     = margin(4, 6, 4, 6)
    )
  
  list(
    delta   = p_delta,
    profile = p_prof
  )
}

load("no_seasonality_gumbel.RData")

plots <- coverage_plots(noseasonality100_gumbel$raw, noseasonality100_gumbel$summary$true_rl)

plots$delta 
plots$profile

#######################################################################

raw <- noseasonality100_gumbel$raw
true_rl <- noseasonality100_gumbel$summary$true_rl


box_df <- raw %>%
  select(delta_rl, prof_rl) %>%   
  pivot_longer(cols = everything(),
               names_to  = "method",
               values_to = "rl") %>%
  mutate(method = recode(method,
                          delta_rl = "Delta method",
                          prof_rl  = "Profile likelihood"))

p_box <- ggplot(box_df, aes(x = method, y = rl, fill = method)) +
  geom_boxplot(width = 0.4, alpha = 0.7, outlier.shape = 16, outlier.alpha = 0.4) +
  geom_hline(yintercept = true_rl, linewidth = 0.5, colour = "black") +
  annotate("text", x = 0.55, y = true_rl,
           label = "True Return Level", hjust = 0.3, vjust = -0.5,
           size = 3.2, colour = "black") +
  scale_fill_manual(values = c("Delta method" = "lightblue",
                                "Profile likelihood" = "steelblue")) +
  
  labs(x = NULL, 
       y = "Return Level") +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.title         = element_text(face = "bold", hjust = 0.5),
    legend.position    = "none"
  )

p_box



```

## Overlay plot to find the sliver 
```{r Overlay plots}

x_samples <- function(n_samples, mu, sigma) {
  month_days  <- c(31,28,31,30,31,30,31,31,30,31,30,31)
  month_index <- rep(1:12, times = month_days)
  month_index <- rep(month_index, n_samples/365)
  rnorm(length(month_index), mean = mu[month_index], sd = sigma[month_index])
}

set.seed(123)

samples <- 1000 * 365
mu <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma <- rep(1,12)

x <- x_samples(samples, mu, sigma)

plot(x[1:(5*365)])

years <- length(x) / 365

plot_df <- data.frame(
  Year = rep(1:years, each = 365),
  Day = rep(1:365, years),
  Value = x
)

# Find annual maxima
min_day <- min(max_idx$Day)
max_day <- max(max_idx$Day)

ind_max <- ggplot(plot_df, aes(Day, Value)) +
  geom_line(
    colour = "lightblue3",
    alpha = 0.4,
    linewidth = 0.6
  ) +
  geom_point(
    data = max_idx,
    aes(Day, Value),
    colour = "black",
    size = 1.8,
    alpha = 0.5
  ) +
  geom_vline(
    xintercept = min_day,
    colour = "black",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  geom_vline(
    xintercept = max_day,
    colour = "black",
    linetype = "dashed",
    linewidth = 0.8
  ) +
  labs(
    x = "Day of the year",
    y = "Observation"
  ) +
  theme_bw()

ind_max

```

## Improved version to match the style with the whole paper and also a condensed one for the other 4 
```{r Better Overlay plots}

x_samples <- function(n_samples, mu, sigma) {
  month_days  <- c(31,28,31,30,31,30,31,31,30,31,30,31)
  month_index <- rep(1:12, times = month_days)
  month_index <- rep(month_index, n_samples/365)
  rnorm(length(month_index), mean = mu[month_index], sd = sigma[month_index])
}

set.seed(123)

samples <- 1000 * 365
mu      <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma   <- rep(1,12)

x     <- x_samples(samples, mu, sigma)
years <- length(x) / 365

plot_df <- data.frame(
  Year  = rep(1:years, each = 365),
  Day   = rep(1:365, years),
  Value = x
)

n_show <- 150                                  
keep   <- sort(sample(years, n_show))

bg_df <- plot_df[plot_df$Year %in% keep, ]

# Annual max 
X       <- matrix(x, nrow = 365, ncol = years)
max_day <- apply(X, 2, which.max)
max_val <- apply(X, 2, max)

max_idx <- data.frame(Year = 1:years, Day = max_day, Value = max_val)

# Effective block size 
pi_t  <- tabulate(max_day, nbins = 365) / years
n_eff <- 1 / sum(pi_t^2)

cat("n_eff =", round(n_eff, 1),
    "| maxima span days", min(max_day), "to", max(max_day), "\n")

# Panels 
p_hist <- ggplot(max_idx, aes(Day)) +
  geom_histogram(binwidth = 7, fill = "grey40") +
  scale_x_continuous(limits = c(1, 365), expand = c(0.01, 0)) +
  labs(x = NULL, y = "Count") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text.x  = element_blank(),
        axis.ticks.x = element_blank())

ind_max <- ggplot() +
  geom_point(data = bg_df, aes(Day, Value, group = Year),
            colour = "lightblue", alpha = 0.2, size = 1) +
  geom_point(data = max_idx, aes(Day, Value),
             colour = "black", size = 1.5, alpha = 1.8) +
  geom_vline(xintercept = range(max_day),
             colour = "black", linetype = "dashed", linewidth = 1) +
  scale_x_continuous(limits = c(1, 365), expand = c(0.01, 0)) +
  labs(x = "Day of the year", y = "Observation") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

p_hist / ind_max + plot_layout(heights = c(1, 5))

ind_max


```

## This below is a small scrap from my 4th meeting with Emma 

Overall from this meeting:

+ It is not supposed to be set $\xi$ to 0 when you think it is gumbel, make the if function for whatever that happens 
+ Would putting the not 0 case first before everything else be faster than putting the if 0 condition?
+ This section is simply just for the case of reformatting our previous case so that it is correct 

+ Looking at the Frechet case, could instead use a linear mean term that would be the mean for each month 
+ STOP looking at values of $\xi<-0.5$ as it is not a very good method of 
+ For the case which looks like the red sea data, take averages of each month for each corresponding year then use that to find averages for the corresponding $\mu$ and $\sigma$ that can be used for simulating samples 
+ Going to sample 50 cases but set it so that it can be changed back to 16 to compare with the red sea data case 
+ What is the result of producing these 
+ Generally just fix the cases and produce an autonomous case for the different cases here 
+ Try the version for a non seasonal case 
+ Make sure the coverage is about 95% for a profile case 

```{r Simple one sample case here }

GEV_samples <- 50 * 365
true_samples <- 100000 * 365


x_samples <- function(n_samples, mu, sigma) {
  
  # holder
  x <- numeric(n_samples)
  
  month_days <- c(31,28,31,30,31,30,31,31,30,31,30,31)

  month_index <- rep(1:12, times = month_days)
  month_index <- rep(month_index, n_samples/365) 
  
  x_gumbel <- rnorm(length(month_index), mean = mu[month_index], sd = sigma[month_index])
  
  x_frechet <- mu[month_index] + rt(length(month_index), df = 5) 
  
  x_weibull <- rbeta(length(month_index), mu[month_index], 10*sigma)
  
  # Make this dataframe so its easier to pull 
  samples <- data.frame(
    gumbel = x_gumbel,
    frechet = x_frechet,
    weibull = x_weibull
  )
  
  samples
  
}

set.seed(123)

# mean just increments 1 to 6 and back down again like a sine wave i guess
mu <- c(1,2,3,4,5,6,6,5,4,3,2,1)
sigma <- c(1,1,1,1,1,1,1,1,1,1,1,1)

x_gev <- x_samples(GEV_samples, mu, sigma)

plot(x_gev$gumbel[1:1000])
plot(x_gev$frechet[1:1000])
plot(x_gev$weibull[1:1000])

block_max <- function(x, block_size) {
  maxima <- tapply(x, (seq_along(x) - 1) %/% block_size, max)
  maxima <- as.numeric(maxima)
}

bm_gumbel <- block_max(x_gev$gumbel, 365)
bm_frechet <- block_max(x_gev$frechet,365)
bm_weibull <- block_max(x_gev$weibull, 365)

#plot(bm_gumbel)
#plot(bm_frechet)
#plot(bm_weibull)

fit_gumbel <- gev.fit(bm_gumbel)
fit_frechet <- gev.fit(bm_frechet)
fit_weibull <- gev.fit(bm_weibull) 

gev.diag(fit_gumbel)
gev.diag(fit_frechet)
gev.diag(fit_weibull)

# Literally do not really understand what to do here and how to use this
# gev.prof(z = fit_gumbel, m = 100, xlow = 80, xup = 120, conf = 0.95, nint = 1000)

# Check the quantiles for the return levels to see if they match up 
x_estimate <- x_samples(true_samples, mu, sigma)

bm_estimate_gumbel  <- block_max(x_estimate$gumbel, 365)
bm_estimate_frechet <- block_max(x_estimate$frechet, 365)
bm_estimate_weibull <- block_max(x_estimate$weibull, 365)

# The quantiles required
T <- c(10,20,100,1000)
quantiles <- 1 - 1/T 

true_rl_gumbel <- data.frame(
  ReturnPeriod = T,
  True = sapply(T, function(t) {
    quantile(bm_estimate_gumbel, probs = 1 - 1/t)
  })
)

true_rl_frechet <- data.frame(
  ReturnPeriod = T,
  True = sapply(T, function(t) {
    quantile(bm_estimate_frechet, probs = 1 - 1/t)
  })
)

true_rl_weibull <- data.frame(
  ReturnPeriod = T,
  True = sapply(T, function(t) {
    quantile(bm_estimate_weibull, probs = 1 - 1/t)
  })
)

# made it so the rl and ci are both in the same dataframe
delta_rl_CI <- function(fit, T) {

  # Extract MLEs
  mu    <- fit$mle[1]
  sigma <- fit$mle[2]
  xi    <- fit$mle[3]

  # Covariance matrix of the estimates
  V <- fit$cov

  y <- -log(1 - 1/T)

  if (abs(xi) < 1e-6) {

    rl <- mu - sigma * log(y)

    grad_rl <- c(1,-log(y))

    var_rl <- V[1,1] + (log(y))^2 * V[2,2] - 2 * log(y) * V[1,2]

  } else {

    rl <- mu + (sigma/xi) * (y^(-xi) - 1)

    grad_rl <- c(1,(y^(-xi) - 1)/xi, sigma * (1 - y^(-xi) - xi * y^(-xi) * log(y))/xi^2)

    var_rl <- as.numeric(t(grad_rl) %*% V %*% grad_rl)
  }

  se_rl <- sqrt(var_rl)

  data.frame(
    ReturnPeriod = T,
    ReturnLevel  = rl,
    Lower        = rl - 1.96 * se_rl,
    Upper        = rl + 1.96 * se_rl
  )
}

est_rl_gumbel <- do.call(rbind, lapply(T, delta_rl_CI, fit = fit_gumbel))
est_rl_frechet <- do.call(rbind, lapply(T, delta_rl_CI, fit = fit_frechet))
est_rl_weibull <- do.call(rbind, lapply(T, delta_rl_CI, fit = fit_weibull))

est_rl_gumbel$True <- quantile(bm_estimate_gumbel, quantiles)
est_rl_frechet$True <- quantile(bm_estimate_frechet, quantiles)
est_rl_weibull$True <- quantile(bm_estimate_weibull, quantiles)

est_rl_gumbel
est_rl_frechet
est_rl_weibull
```

