---
title: "Red Sea Missingness"
author: '25062476'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ismev) # Coles Package
library(evmissing) # Simpson & Northrop package 
library(readr) # for csv and things 
library(ggplot2)
library(dplyr)
library(patchwork)

library(kableExtra)

set.seed(123)

load("Red_Sea_weekly.RData")
```

# Note that this is literally just for one case 
One singular 50 year sample here 
Testing here for all the cases that I would need 

```{r Functions and pre run}
x_samples_weekly <- function(n_samples, mu, sigma) {
  week_days  <- rep(7, 51)
  week_days  <- c(week_days, 8)
  week_index <- rep(1:52, times = week_days)
  week_index <- rep(week_index, n_samples / 365)
  rnorm(length(week_index), mean = mu[week_index], sd = sigma[week_index])
}

block_max <- function(x, block_size) {
  as.numeric(tapply(x, (seq_along(x) - 1) %/% block_size, max))
}
 
fit_and_extract <- function(x, block_length, T, discard = NULL, adjust = FALSE, min_blocks = 10) {
 
  fit <- tryCatch(
    suppressWarnings(
      if (is.null(discard)) {
        gev_mle(data = x, block_length = block_length, adjust = adjust)
      } else {
        gev_mle(data = x, block_length = block_length, adjust = adjust, discard = discard)
      }
    ),
    error = function(e) NULL
  )
 
  n_blocks <- if (!is.null(fit)) length(fit$maxima) else 0L
 
  if (is.null(fit) || fit$convergence != 0 || length(fit$par) != 3 ||
      any(is.na(fit$par)) || n_blocks < min_blocks) {
    return(c(rl = NA_real_, lower = NA_real_, upper = NA_real_, n_blocks = n_blocks))
  }
 
  ret <- tryCatch(gev_return(fit, m = T), error = function(e) NULL)
  if (is.null(ret)) {
    return(c(rl = NA_real_, lower = NA_real_, upper = NA_real_, n_blocks = n_blocks))
  }
  rl <- as.numeric(ret)
 
  ci <- tryCatch(
    suppressWarnings(confint(ret, profile = TRUE)),
    error = function(e) NULL
  )
 
  if (is.null(ci)) {
    return(c(rl = rl, lower = NA_real_, upper = NA_real_, n_blocks = n_blocks))
  }
 
  out <- c(rl = rl, lower = as.numeric(ci[1, 1]), upper = as.numeric(ci[1, 2]), n_blocks = n_blocks)
  if (length(out) != 4) {
    return(c(rl = NA_real_, lower = NA_real_, upper = NA_real_, n_blocks = n_blocks))
  }
  out
}
 
missing_prof_week <- function(GEV_samples, n_samples, q_samples, T, mu, sigma, loc_id,
                               p_max = 0.2, discard_thresh = 15, verbose = TRUE) {
 
  x_true  <- x_samples_weekly(q_samples, mu, sigma)
  bm_true <- block_max(x_true, 365)
  true_rl <- as.numeric(quantile(bm_true, probs = 1 - 1 / T))
 
  n_years <- GEV_samples / 365
  methods <- c("full", "naive", "discard", "adjust")
 
  col_names <- unlist(lapply(methods, function(m) paste0(m, c("_rl", "_lower", "_upper", "_nblocks"))))
  results <- as.data.frame(matrix(NA_real_, nrow = n_samples, ncol = length(col_names)))
  names(results) <- col_names
 
  fail_counts <- setNames(rep(0L, length(methods)), methods)
 
  for (i in seq_len(n_samples)) {
 
    if (verbose && i %% max(1, n_samples %/% 20) == 0)
      cat(round(100 * i / n_samples), "% done\n")
 
    x_full <- x_samples_weekly(GEV_samples, mu, sigma)
 
    years <- split(x_full, rep(seq_len(n_years), each = 365))
    
    for (yr in seq_len(n_years)) {
      p <- runif(1, 0, p_max)
      n_remove <- floor(p * 365)
      if (n_remove > 0) {
        miss <- sample(365, n_remove)
        years[[yr]][miss] <- NA
      }
    }
    x_missing <- unlist(years)
 
    out <- fit_and_extract(x_full, block_length = 365, T = T)
    results[i, c("full_rl", "full_lower", "full_upper", "full_nblocks")] <- out
    if (is.na(out["rl"])) fail_counts["full"] <- fail_counts["full"] + 1L
 
    out <- fit_and_extract(x_missing, block_length = 365, T = T, adjust = FALSE)
    results[i, c("naive_rl", "naive_lower", "naive_upper", "naive_nblocks")] <- out
    if (is.na(out["rl"])) fail_counts["naive"] <- fail_counts["naive"] + 1L
 
    out <- fit_and_extract(x_missing, block_length = 365, T = T,
                            discard = discard_thresh, adjust = FALSE)
    results[i, c("discard_rl", "discard_lower", "discard_upper", "discard_nblocks")] <- out
    if (is.na(out["rl"])) fail_counts["discard"] <- fail_counts["discard"] + 1L
 
    out <- fit_and_extract(x_missing, block_length = 365, T = T, adjust = TRUE)
    results[i, c("adjust_rl", "adjust_lower", "adjust_upper", "adjust_nblocks")] <- out
    if (is.na(out["rl"])) fail_counts["adjust"] <- fail_counts["adjust"] + 1L
  }
 
  summarize_method <- function(prefix) {
    rl      <- results[[paste0(prefix, "_rl")]]
    lower   <- results[[paste0(prefix, "_lower")]]
    upper   <- results[[paste0(prefix, "_upper")]]
    nblocks <- results[[paste0(prefix, "_nblocks")]]
 
    data.frame(
      method       = prefix,
      ReturnPeriod = T,
      true_rl      = true_rl,
      mean_rl      = mean(rl, na.rm = TRUE),
      Bias_mean    = mean(rl - true_rl, na.rm = TRUE),
      Bias_median  = median(rl - true_rl, na.rm = TRUE),
      SD           = sd(rl, na.rm = TRUE),
      RMSE         = sqrt(mean((rl - true_rl)^2, na.rm = TRUE)),
      Coverage     = mean(lower <= true_rl & upper >= true_rl, na.rm = TRUE),
      CI_width     = mean(upper - lower, na.rm = TRUE),
      mean_n_blocks = mean(nblocks, na.rm = TRUE),
      n_fail       = fail_counts[[prefix]],
      n_valid      = sum(!is.na(rl))
    )
  }
 
  summary_df <- do.call(rbind, lapply(methods, summarize_method))
  rownames(summary_df) <- NULL
 
  list(
    summary     = summary_df,
    raw         = results,
    true_rl     = true_rl,
    fail_counts = fail_counts
  )
}

load("Red_Sea_weekly.RData")

```

# Location 1
```{r To load and run the data }

loc_id      <- 1
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

set.seed(123+loc_id) 
missing_loc1_rs25 <- missing_prof_week(GEV_samples, n_samples, q_samples, 25, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc1_rs50 <- missing_prof_week(GEV_samples, n_samples, q_samples, 50, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc1_rs100 <- missing_prof_week(GEV_samples, n_samples, q_samples, 100, mu, sigma, loc_id)

print(missing_loc1_rs25$summary)
print(missing_loc1_rs50$summary)
print(missing_loc1_rs100$summary)

save(missing_loc1_rs25, 
     missing_loc1_rs50,
     missing_loc1_rs100,
     file = "missing_loc1_rs.RData")

```

# Location 2 
```{r Second location}

loc_id      <- 2
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

set.seed(123+loc_id) 
missing_loc2_rs25 <- missing_prof_week(GEV_samples, n_samples, q_samples, 25, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc2_rs50 <- missing_prof_week(GEV_samples, n_samples, q_samples, 50, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc2_rs100 <- missing_prof_week(GEV_samples, n_samples, q_samples, 100, mu, sigma, loc_id)

print(missing_loc2_rs25$summary)
print(missing_loc2_rs50$summary)
print(missing_loc2_rs100$summary)

save(missing_loc2_rs25, 
     missing_loc2_rs50,
     missing_loc2_rs100,
     file = "missing_loc2_rs.RData")

```

# Location 3 
```{r Third location}

loc_id      <- 3
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

set.seed(123+loc_id) 
missing_loc3_rs25 <- missing_prof_week(GEV_samples, n_samples, q_samples, 25, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc3_rs50 <- missing_prof_week(GEV_samples, n_samples, q_samples, 50, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc3_rs100 <- missing_prof_week(GEV_samples, n_samples, q_samples, 100, mu, sigma, loc_id)

print(missing_loc3_rs25$summary)
print(missing_loc3_rs50$summary)
print(missing_loc3_rs100$summary)

save(missing_loc3_rs25, 
     missing_loc3_rs50,
     missing_loc3_rs100,
     file = "missing_loc3_rs.RData")

```

# Location 4 
```{r Fourth location}

loc_id      <- 4
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

set.seed(123+loc_id) 
missing_loc4_rs25 <- missing_prof_week(GEV_samples, n_samples, q_samples, 25, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc4_rs50 <- missing_prof_week(GEV_samples, n_samples, q_samples, 50, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc4_rs100 <- missing_prof_week(GEV_samples, n_samples, q_samples, 100, mu, sigma, loc_id)

print(missing_loc4_rs25$summary)
print(missing_loc4_rs50$summary)
print(missing_loc4_rs100$summary)

save(missing_loc4_rs25, 
     missing_loc4_rs50,
     missing_loc4_rs100,
     file = "missing_loc4_rs.RData")

```

# Location 5 
```{r Fifth location}

loc_id      <- 5
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

set.seed(123+loc_id) 
missing_loc5_rs25 <- missing_prof_week(GEV_samples, n_samples, q_samples, 25, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc5_rs50 <- missing_prof_week(GEV_samples, n_samples, q_samples, 50, mu, sigma, loc_id)
set.seed(123+loc_id) 
missing_loc5_rs100 <- missing_prof_week(GEV_samples, n_samples, q_samples, 100, mu, sigma, loc_id)

print(missing_loc5_rs25$summary)
print(missing_loc5_rs50$summary)
print(missing_loc5_rs100$summary)

save(missing_loc5_rs25, 
     missing_loc5_rs50,
     missing_loc5_rs100,
     file = "missing_loc5_rs.RData")

```

# Data loading for location 1 to 3 for a table 
Note that this is way too much data to include locations 1 to 5 as the standard 5 versions throughout the whole paper so only 3 are put here then the next 2 locations are put in another table
```{r Loading locations 1 to 3}
load("missing_loc1_rs.RData")
load("missing_loc2_rs.RData")
load("missing_loc3_rs.RData")
load("missing_loc4_rs.RData")
load("missing_loc5_rs.RData")

print(missing_loc1_rs25$summary)
print(missing_loc1_rs50$summary)
print(missing_loc1_rs100$summary)

print(missing_loc2_rs25$summary)
print(missing_loc2_rs50$summary)
print(missing_loc2_rs100$summary)

print(missing_loc3_rs25$summary)
print(missing_loc3_rs50$summary)
print(missing_loc3_rs100$summary)

print(missing_loc4_rs25$summary)
print(missing_loc4_rs50$summary)
print(missing_loc4_rs100$summary)

print(missing_loc5_rs25$summary)
print(missing_loc5_rs50$summary)
print(missing_loc5_rs100$summary)
```

# Histogram of the return level histograms for the different methods
```{r Producing histogram of the return levels and how the mean and median bias differs}
load("missing_loc1_rs.RData")

true_rl           <- missing_loc1_rs100$true_rl
full_samples      <- missing_loc1_rs100$raw[, 1]
naive_samples     <- missing_loc1_rs100$raw[, 5]
discard_samples   <- missing_loc1_rs100$raw[, 9]
adjust_samples    <- missing_loc1_rs100$raw[, 13]

all_vals <- c(full_samples, naive_samples, discard_samples, adjust_samples)
x_lims   <- c(28.35,29.2)
# range(all_vals, na.rm = TRUE)

make_hist_panel <- function(samples, label, true_rl, x_lims, bins = 30, show_x_title = TRUE) {

  mean_val   <- mean(samples, na.rm = TRUE)
  median_val <- median(samples, na.rm = TRUE)

  df       <- data.frame(x = samples)
  line_df  <- data.frame(
    value = c(true_rl, mean_val, median_val),
    type  = factor(
      c("Ref RL", "Mean Bias", "Median Bias"),
      levels = c("Ref RL", "Mean Bias", "Median Bias")
    )
  )

  ggplot(df, aes(x = x)) +
    geom_histogram(bins = bins, fill = "lightblue", color = "white") +
    geom_vline(
      data = line_df,
      aes(xintercept = value, color = type, linetype = type),
      linewidth = 1
    ) +
    scale_color_manual(
      name   = NULL,
      values = c("Ref RL" = "black", "Mean Bias" = "blue", "Median Bias" = "red"),
      guide  = guide_legend(keywidth = unit(2, "cm"))
    ) +
    scale_linetype_manual(
      name   = NULL,
      values = c("Ref RL" = "solid", "Mean Bias" = "dashed", "Median Bias" = "dotted"),
      guide  = guide_legend(keywidth = unit(2, "cm"))
    ) +
    coord_cartesian(xlim = x_lims) +
    labs(title = label, 
         x = if (show_x_title) "100 year return level" else NULL,
         y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title      = element_text(face = "bold"),
      legend.position = "bottom"
    )
}

p11 <- make_hist_panel(full_samples,    "Full",    true_rl, x_lims, show_x_title = FALSE)
p21 <- make_hist_panel(naive_samples,   "Naive",   true_rl, x_lims, show_x_title = FALSE)
p31 <- make_hist_panel(discard_samples, "Discard", true_rl, x_lims, show_x_title = FALSE)
p41 <- make_hist_panel(adjust_samples,  "Adjust",  true_rl, x_lims, show_x_title = TRUE)

combined_loc1 <- (p11 / p21 / p31 / p41) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom",
      legend.key.height   = unit(0.8, "cm")
      )

combined_loc1

###############################################################################



```

# Block maxima visualisation AND the histogram of GEV fit with the different methods
```{r Block Maxima visualisation plots}

set.seed(123)  

loc_id      <- 1
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

x_full <- x_samples_weekly(GEV_samples, mu, sigma)
n_years <- GEV_samples / 365
years <- split(x_full, rep(seq_len(n_years), each = 365))
for (yr in seq_len(n_years)) {
  p <- runif(1, 0, 0.2)
  n_remove <- floor(p * 365)
  if (n_remove > 0) years[[yr]][sample(365, n_remove)] <- NA
}
x_missing <- unlist(years)

fit_full    <- gev_mle(data = x_full,    block_length = 365, adjust = FALSE)
fit_naive   <- gev_mle(data = x_missing, block_length = 365, adjust = FALSE)
fit_discard <- gev_mle(data = x_missing, block_length = 365, adjust = FALSE, discard = 15)
fit_adjust  <- gev_mle(data = x_missing, block_length = 365, adjust = TRUE)

dgev <- function(x, mu, sigma, xi){
  if (abs(xi) < 1e-8){
    z <- (x - mu) / sigma
    return(exp(-(z + exp(-z))) / sigma)
  }
  y <- 1 + xi * (x - mu) / sigma
  dens <- ifelse(
    y > 0,
    (1/sigma) * y^(-1/xi - 1) * exp(-y^(-1/xi)),
    0
  )
  dens
}

fits <- list(full = fit_full, naive = fit_naive, discard = fit_discard, adjust = fit_adjust)

mu_full    <- fit_full$par["mu"];    sigma_full    <- fit_full$par["sigma"];    xi_full    <- fit_full$par["xi"]
mu_naive   <- fit_naive$par["mu"];   sigma_naive   <- fit_naive$par["sigma"];   xi_naive   <- fit_naive$par["xi"]
mu_discard <- fit_discard$par["mu"]; sigma_discard <- fit_discard$par["sigma"]; xi_discard <- fit_discard$par["xi"]
mu_adjust  <- fit_adjust$par["mu"];  sigma_adjust  <- fit_adjust$par["sigma"];  xi_adjust  <- fit_adjust$par["xi"]

bm_full <- fit_full$maxima
x_grid  <- seq(min(bm_full), max(bm_full), length.out = 500)

curve_all <- rbind(
  data.frame(method = "full",    x = x_grid, density = dgev(x_grid, mu_full,    sigma_full,    xi_full)),
  data.frame(method = "naive",   x = x_grid, density = dgev(x_grid, mu_naive,   sigma_naive,   xi_naive)),
  data.frame(method = "discard", x = x_grid, density = dgev(x_grid, mu_discard, sigma_discard, xi_discard)),
  data.frame(method = "adjust",  x = x_grid, density = dgev(x_grid, mu_adjust,  sigma_adjust,  xi_adjust))
)

method_levels <- c("full", "naive", "discard", "adjust")
curve_all$method <- factor(curve_all$method, levels = method_levels)

points_full <- data.frame(x = bm_full, y = 0)

example_hist <- ggplot() +
  geom_histogram(
    data = data.frame(x = bm_full),
    aes(x, after_stat(density)),
    bins = 9,
    fill = "lightblue",
    colour = "black"
  ) +
  geom_line(
    data = curve_all,
    aes(x, density, colour = method, linetype = method),
    linewidth = 1.1
  ) +
  geom_point(
    data = points_full,
    aes(x = x, y = 0),
    size = 2
  ) +
  scale_colour_manual(values = c(full = "black", naive = "firebrick",
                                  discard = "orange", adjust = "blue")) +
  scale_linetype_manual(values = c(full = "solid", naive = "dashed",
                                    discard = "dashed", adjust = "dashed")) +
  labs(
    x = "Annual maxima",
    y = "Density",
    colour = "Method",
    linetype = "Method"
  ) +
  theme_bw(base_size = 14) +
  theme(
      panel.grid.major = element_blank(),
      plot.title       = element_text(face = "bold", hjust = 0.5),
      legend.position  = c(0.85, 0.75),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin    = margin(4, 6, 4, 6),
      legend.key.width   = unit(1.6, "cm")
    )

example_hist

```

# Double side by side plot of block maxima and the histogram of GEV fit
```{r Same as above but with block maxima}
set.seed(123)  

loc_id      <- 1
GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu          <- week_mean_all[,loc_id]
sigma       <- week_sd_all[,loc_id]

GEV_samples <- 50 * 365
x_full  <- x_samples_weekly(GEV_samples, mu, sigma)
n_years <- GEV_samples / 365

years        <- split(x_full, rep(seq_len(n_years), each = 365))
missing_flag <- rep(FALSE, GEV_samples)

for (yr in seq_len(n_years)) {
  p <- runif(1, 0, 0.2)
  n_remove <- floor(p * 365)
  if (n_remove > 0) {
    miss <- sample(365, n_remove)
    idx_in_full <- (yr - 1) * 365 + miss
    missing_flag[idx_in_full] <- TRUE
  }
}

df <- data.frame(
  Index         = seq_len(GEV_samples),
  Observations  = x_full,
  Missing       = missing_flag
)
df$Observations_missing <- ifelse(df$Missing, NA, df$Observations)

# Block maxima data frame 
block_max_df <- function(df, value_col, block_size, na.rm = TRUE) {
  block <- (df$Index - 1) %/% block_size
  do.call(rbind, lapply(split(df, block), function(d) {
    v <- d[[value_col]]
    if (na.rm) d <- d[!is.na(v), ]
    if (nrow(d) == 0) return(NULL)
    d[which.max(d[[value_col]]), ]
  }))
}

max_points_full    <- block_max_df(df, "Observations",          365, na.rm = FALSE)
max_points_observed <- block_max_df(df, "Observations_missing", 365, na.rm = TRUE)

# Left figure for building the block maxima similar to chapter 2 
bm_full <- ggplot(df, aes(Index, Observations)) +
  geom_point(aes(colour = "Observations"), size = 1.5) +
  geom_vline(xintercept = seq(365, 365 * (n_years - 1), by = 365),
             alpha = 0.5, linetype = "dashed") +
  geom_point(aes(colour = "Annual maxima"), data = max_points_full, size = 3) +
  scale_colour_manual(values = c("Observations" = "lightblue", "Annual maxima" = "black")) +
  labs(x = "Observation Index", y = "Data", colour = NULL, title = "Full data") +
  guides(colour = guide_legend(override.aes = list(size = c(3, 3)))) +
  theme_bw(base_size = 14) +
  theme(
    panel.grid.major   = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.title         = element_text(face = "bold", hjust = 0.5),
    legend.position    = c(0.75, 0.10),
    legend.background  = element_rect(fill = "white", colour = "grey70"),
    legend.margin      = margin(4, 6, 4, 6)
  )

# Right figure for building histogram similar to the histogram in chapter 2 
bm_missing <- ggplot(df, aes(Index, Observations)) +
  geom_point(aes(colour = "Observations"), size = 1.5) +
  geom_point(data = subset(df, Missing), aes(colour = "Missing"), size = 1.5) +
  geom_vline(xintercept = seq(365, 365 * (n_years - 1), by = 365),
             alpha = 0.5, linetype = "dashed") +
  geom_point(aes(colour = "Observed annual max"), data = max_points_observed, size = 3) +
  scale_colour_manual(values = c(
    "Observations"        = "lightblue",
    "Missing"              = "firebrick",
    "Observed annual max"  = "black"
  )) +
  labs(x = "Observation Index", y = "Data", colour = NULL, title = "Missing data mechanism applied") +
  guides(colour = guide_legend(override.aes = list(size = c(3, 1.5, 3)))) +
  theme_bw(base_size = 14) +
  theme(
    panel.grid.major   = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.title         = element_text(face = "bold", hjust = 0.5),
    legend.position    = c(0.75, 0.10),
    legend.background  = element_rect(fill = "white", colour = "grey70"),
    legend.margin      = margin(4, 6, 4, 6)
  )

bm_full + bm_missing
```

# All missing individual simulation tables and the helper function used to make it
```{r All the individual tables}

prepare_table_missing <- function(rs_list, scenario) {
  bind_rows(lapply(rs_list, function(x) x$summary)) |>
    select(
      ReturnPeriod,
      method,
      true_rl,
      mean_rl,
      RMSE,
      Bias_mean,
      Bias_median,
      Coverage,
      CI_width,
      mean_n_blocks
    ) |>
    rename(
      T          = ReturnPeriod,
      Method     = method,
      `True RL`  = true_rl,
      `Est RL`   = mean_rl,
      Bias       = Bias_mean,
      `MD Bias`  = Bias_median,
      Width      = CI_width,
      `N Blocks` = mean_n_blocks
    ) |>
    mutate(
      Scenario = scenario,
      Method = factor(Method, levels = c("full", "naive", "discard", "adjust")),
      .before = 1
    ) |>
    arrange(T, Method)
}

###############################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc1_rs25, missing_loc1_rs50, missing_loc1_rs100), "RS1"),
  prepare_table_missing(list(missing_loc2_rs25, missing_loc2_rs50, missing_loc2_rs100), "RS2"),
  prepare_table_missing(list(missing_loc3_rs25, missing_loc3_rs50, missing_loc3_rs100), "RS3")
)

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 1 to 3. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdatafull"
  ) |>
  save_kable("missingdataloc123.tex")

################################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc4_rs25, missing_loc4_rs50, missing_loc4_rs100), "RS4"),
  prepare_table_missing(list(missing_loc5_rs25, missing_loc5_rs50, missing_loc5_rs100), "RS5")
)

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 4 and 5. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdata45"
  ) |>
  save_kable("missingdataloc45.tex")

################################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc1_rs25, missing_loc1_rs50, missing_loc1_rs100), "RS1")
)

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 1. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdatafull"
  ) |>
  save_kable("missingdataloc1.tex")

################################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc2_rs25, missing_loc2_rs50, missing_loc2_rs100), "RS2")
)

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 2. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdatafull"
  ) |>
  save_kable("missingdataloc2.tex")

################################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc3_rs25, missing_loc3_rs50, missing_loc3_rs100), "RS3")
)

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 3. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdatafull"
  ) |>
  save_kable("missingdataloc3.tex")

################################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc4_rs25, missing_loc4_rs50, missing_loc4_rs100), "RS4")
  )

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 4. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdatafull"
  ) |>
  save_kable("missingdataloc4.tex")

################################################################################

missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc5_rs25, missing_loc5_rs50, missing_loc5_rs100), "RS5")
)

missing_all <- missing_all |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))

missing_all

missing_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Return level estimation performance for the weekly-seasonal Red Sea missing data simulation Locations 5. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:missingdatafull"
  ) |>
  save_kable("missingdataloc5.tex")




```

# Making it into a singular fat table 
```{r Putting it into one }
missing_all <- bind_rows(
  prepare_table_missing(list(missing_loc1_rs25, missing_loc1_rs50, missing_loc1_rs100), "RS1"),
  prepare_table_missing(list(missing_loc2_rs25, missing_loc2_rs50, missing_loc2_rs100), "RS2"),
  prepare_table_missing(list(missing_loc3_rs25, missing_loc3_rs50, missing_loc3_rs100), "RS3"),
  prepare_table_missing(list(missing_loc4_rs25, missing_loc4_rs50, missing_loc4_rs100), "RS4"),
  prepare_table_missing(list(missing_loc5_rs25, missing_loc5_rs50, missing_loc5_rs100), "RS5")
)

```

# Junkyard
This is a bit of older stuff that I took my own time to just make without realising that the functions could have already been done by evmissing package

```{r Testing functions}

x_samples_weekly <- function(n_samples, mu, sigma) {
  week_days  <- rep(7,51)
  week_days <- c(week_days,8)
  week_index <- rep(1:52, times = week_days)
  week_index <- rep(week_index, n_samples/365)
  rnorm(length(week_index), mean = mu[week_index], sd = sigma[week_index])
}

block_max <- function(x, block_size) {
  as.numeric(
    tapply(x,(seq_along(x) - 1) %/% block_size,max,na.rm = TRUE)
  )
}

block_max_discard <- function(x, block_size, threshold = 0.15) {

  f <- function(z){
    prop_missing <- mean(is.na(z))
    
    if(prop_missing > threshold){
      return(NA_real_)
    }
    max(z, na.rm = TRUE)
  }

  as.numeric(
    tapply(
      x,
      (seq_along(x)-1) %/% block_size,
      f
    )
  )
}

load("Red_Sea_weekly.RData")

set.seed(123)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu <- week_mean_all[,1]
sigma <- week_sd_all[,1]

x_truth <- x_samples_weekly(q_samples, mu, sigma)

x <- x_samples_weekly(GEV_samples, mu, sigma)

y <- -log(1 - 1/100) # Note this 100 return level for all 

years <- split(x, rep(1:50, each = 365))

for(i in 1:50){

    p <- runif(1, 0, 0.2)

    n_remove <- floor(p * 365)

    miss <- sample(365, n_remove)

    years[[i]][miss] <- NA
}

x_missing <- unlist(years)

################################## full version

set.seed(123)

bm_full <- block_max(x_truth,365)
fit_full <- gev.fit(bm_full)

# do not really need to see this diagnostic as it would just be hella perfect 

#################################### naive version WITH DELTA ISMEV

naive_bm <- block_max(x_missing,365)
# plot(naive_bm)
fit_naive <- gev.fit(naive_bm)
gev.diag(fit_naive)

#################################### discard version WITH DELTA ISMEV

discard_bm <- block_max_discard(x_missing,365)
discard_bm <- na.omit(discard_bm)
# plot(discard_bm)
fit_discard <- gev.fit(discard_bm)
gev.diag(fit_discard)

#################################### naive version with profile evmissing

fit_naive_prof <- gev_mle(data = x_missing, 
                      block_length = 365, 
                      adjust = FALSE)

plot(fit_naive_prof)

#################################### discard version with profile evmissing

fit_discard_prof <- gev_mle(data = x_missing, 
                      block_length = 365, 
                      adjust = FALSE,
                      discard = 15)

plot(fit_discard_prof)

#################################### Adjust method for profile evmissing

fit_adjust <- gev_mle(data = x_missing, 
                      block_length = 365, 
                      adjust = TRUE)

plot(fit_adjust)


####################################### 100 year rl for 
# full, naive delta, discard delta, naive prof, discard prof, adjust prof

fit_full$mle[1] + (fit_full$mle[2]/fit_full$mle[3]) * (y^(-fit_full$mle[3]) - 1)
fit_naive$mle[1] + (fit_naive$mle[2]/fit_naive$mle[3]) * (y^(-fit_naive$mle[3]) - 1)
fit_discard$mle[1] + (fit_discard$mle[2]/fit_discard$mle[3]) * (y^(-fit_discard$mle[3]) - 1)
gev_return(fit_naive_prof)
gev_return(fit_discard_prof)
gev_return(fit_adjust)






```

```{r Attempt at the fat function}

missing_prof_week <- function(GEV_samples, n_samples, q_samples, T, mu, sigma) {
  
  # For the true 100000 samples that we take as the quantile
  x_true <- x_samples_weekly(q_samples, mu, sigma)
  bm_true <- block_max(x_true, 365)
  true_rl <- as.numeric(quantile(bm_true, probs = 1 - 1/T))
  
  # Results storage — one row per simulation
  results <- data.frame(
    full_rl     = numeric(n_samples),
    full_lower  = numeric(n_samples),
    full_upper  = numeric(n_samples),
    
    naive_rl     = numeric(n_samples),
    naive_lower  = numeric(n_samples),
    naive_upper  = numeric(n_samples),
    
    discard_rl     = numeric(n_samples),
    discard_lower  = numeric(n_samples),
    discard_upper  = numeric(n_samples),
    
    adjust_rl     = numeric(n_samples),
    adjust_lower  = numeric(n_samples),
    adjust_upper  = numeric(n_samples)
  )
  
  for (i in seq_len(n_samples)) {

    if (i %% max(1, n_samples %/% 4) == 0)
      cat(round(100 * i / n_samples), "% done\n")
    
    x_full <- x_samples_weekly(GEV_samples, mu, sigma) # for full

    y <- -log(1 - 1/100) # Note this 100 return level for all 
    
    years <- split(x_full, rep(1:50, each = 365))
    
    for(i in 1:50){
        p <- runif(1, 0, 0.2)
        n_remove <- floor(p * 365)
        miss <- sample(365, n_remove)
        years[[i]][miss] <- NA
    }
    
    x_missing <- unlist(years) # use this for naive, discard, and adjust

    # Profile likelihood
    fit_full <- tryCatch(
      suppressWarnings(gev_mle(x_full, 1)),
      error = function(e) NULL
    )

    if (is.null(fit_prof) || any(is.na(fit_prof))) {
      profile_fail <- profile_fail + 1L
      results[i, c("full_rl", "full_lower", "full_upper")] <- NA
    } else {
      full_return <- gev_return(fit_full, m = T)
      full_conf   <- tryCatch(
        suppressWarnings(confint(prof_return, profile = TRUE)),
        error = function(e) NULL
      )

      if (is.null(prof_conf)) {
        profileci_fail <- profileci_fail + 1L
        results[i, c("full_rl", "full_lower", "full_upper")] <- NA
      } else {
        results[i, "full_rl"]    <- as.numeric(full_return)
        results[i, "full_lower"] <- full_conf[1, 1]
        results[i, "full_upper"] <- full_conf[1, 2]
      }
    }
  }
  
  # Summary
  summary_df <- data.frame(
    ReturnPeriod = T,
    true_rl      = true_rl,
    
    full_rl        = mean(results$full_rl,   na.rm = TRUE),
    full_RMSE      = sqrt(mean((results$full_rl  - true_rl)^2, na.rm = TRUE)),
    full_Bias      = mean(results$full_rl  - true_rl, na.rm = TRUE),
    full_Bias_med  = median(results$full_rl  - true_rl, na.rm = TRUE),
    full_Coverage  = mean(results$full_lower  <= true_rl & results$full_upper  >= true_rl, na.rm = TRUE),
    full_width     = mean(results$full_upper - results$full_lower, na.rm=TRUE)
    
    
    
  )

  list(
    summary            = summary_df,
    raw                = results,        # full simulation results if you need them
    true_rl            = true_rl,
  )
  
}

```

```{r 50 years for location one}
################################################################################
# For the sake of visibility I am going to write out all the things needed for computing this case again

load("Red_Sea_weekly.RData")

set.seed(123)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365
mu <- week_mean_all[,1]
sigma <- week_sd_all[,1]

# Note that this is a bit inefficient but this generates every single one individually and to save it as well 
# Doing the combination of all these functions took way too long and I think this is simply faster if you want a specific return level year

set.seed(123)
test50_redsea_25 <- delta_prof_week(GEV_samples, n_samples, q_samples, 25, mu, sigma)
set.seed(123)
test50_redsea_50 <- delta_prof_week(GEV_samples, n_samples, q_samples, 50, mu, sigma)
set.seed(123)
test50_redsea_100 <- delta_prof_week(GEV_samples, n_samples, q_samples, 100, mu, sigma)

save(test50_redsea_25, 
     test50_redsea_50,
     test50_redsea_100,
     file = "Test_run50_Red_Sea.RData")

```

```{r Test case 50 years Data}
load("Test_run50_Red_Sea.RData")

test50_redsea <- rbind(
  test50_redsea_25$summary, 
  test50_redsea_50$summary,
  test50_redsea_100$summary
)

test50_redsea

plots <- coverage_plots(
  test50_redsea_100$raw, 
  test50_redsea_100$summary$true_rl
  )

plots$delta
plots$profile
```


