---
title: "Red Sea Wave"
author: '25062476'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ismev) # Coles Package
library(evmissing) # Simpson & Northrop package 
library(readr) # for csv and things 
library(ggplot2)
library(dplyr)

set.seed(123)

load("Red_Sea_data.RData")
load("Red_Sea_weekly.RData")
```

# This markdown documnt is generally for producing code for a harmonic function 
Some parts are flipped: e.g. the data is produced from a chunk below then the residual plots of it are plotted in the chunk here:
```{r residual plot of all values here}

load("Red_Sea_data.RData")

x <- data[, 1]

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

plot_df <- data.frame(
  Day = day,
  SST = x,
  Seasonal = seasonal_mean
)

doy_plot <- ggplot(plot_df, aes(doy, SST)) +
  geom_point(alpha = 0.3, colour = "lightblue") +
  geom_line(
    aes(y = Seasonal),
    linewidth = 1,
    colour = "blue"
  ) +
  labs(
    x = "Day of year",
    y = "Sea Surface Temperature (°C)"
  ) +
  theme_bw()

doy_plot

resid_df <- data.frame(
  day   = day,
  doy   = doy,
  fitted = seasonal_mean,
  r      = resid(fit_sin)
  ) %>%
  mutate(month = cut(doy, breaks = c(0, cumsum(c(31,28,31,30,31,30,31,31,30,31,30,31))),
                     labels = month.abb))

p_mean <- ggplot(resid_df, aes(doy, r)) +
  geom_point(alpha = 1, colour = "lightblue", size = 1.2) +
  geom_hline(yintercept = 0, colour = "grey40") +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cc"), linewidth = 1.5 ,colour = "steelblue4", se = FALSE) +
  labs(x = "Day of year", y = "Residual") +
  theme_bw() + theme(panel.grid.minor = element_blank())

p_mean
```

# Data cleaning to see how the smooth harmonic model could fit here
```{r cleaning and plots}

load("Red_Sea_data.RData")

x <- data[, 1]

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

plot_df <- data.frame(
  Day = day,
  SST = x,
  Seasonal = seasonal_mean
)

smooth_rs <- ggplot(plot_df, aes(Day, SST)) +
  geom_line(alpha = 0.4, colour = "lightblue") +
  geom_line(
    aes(y = Seasonal),
    linewidth = 1,
    colour = "blue"
  ) +
  labs(
    x = "Day",
    y = "Sea Surface Temperature (°C)"
  ) +
  theme_bw()

smooth_rs

smooth_5yrs <- ggplot(plot_df[1:(5*365),], aes(Day, SST)) +
  geom_line(alpha = 0.4, colour = "lightblue") +
  geom_line(
    aes(y = Seasonal),
    linewidth = 1,
    colour = "blue"
  ) +
  labs(
    x = "Day",
    y = "Sea Surface Temperature (°C)"
  ) +
  theme_bw()

smooth_5yrs

doy_plot <- ggplot(plot_df, aes(doy, SST)) +
  geom_point(alpha = 0.3, colour = "lightblue") +
  geom_line(
    aes(y = Seasonal),
    linewidth = 1,
    colour = "blue"
  ) +
  labs(
    x = "Day of year",
    y = "Sea Surface Temperature (°C)"
  ) +
  theme_bw()

doy_plot

```

# Adjusted code to sample from the smooth method 
```{r Functions and initialisation}

x_samples_sinusoidal <- function(n_samples, fit, residual_sd) {

  day <- 1:n_samples

  doy <- ((day - 1) %% 365) + 1

  seasonal_mean <- predict(
    fit,
    newdata = data.frame(doy = doy)
  )

  seasonal_mean +
    rnorm(
      n_samples,
      mean = 0,
      sd = residual_sd
    )
}

block_max <- function(x, block_size) {
  as.numeric(tapply(x, (seq_along(x) - 1) %/% block_size, max))
}

delta_rl_CI <- function(T, fit) {
  mu    <- fit$mle[1]
  sigma <- fit$mle[2]
  xi    <- fit$mle[3]
  V     <- fit$cov
  y     <- -log(1 - 1/T)

  if (abs(xi) < 1e-6) {
    rl     <- mu - sigma * log(y)
    var_rl <- V[1,1] + (log(y))^2 * V[2,2] - 2 * log(y) * V[1,2]
  } else {
    rl      <- mu + (sigma/xi) * (y^(-xi) - 1)
    grad_rl <- c(1, (y^(-xi) - 1)/xi, sigma * (1 - y^(-xi) - xi * y^(-xi) * log(y))/xi^2)
    var_rl  <- as.numeric(t(grad_rl) %*% V %*% grad_rl)
  }

  se_rl <- sqrt(var_rl)
  data.frame(ReturnLevel = rl, Lower = rl - 1.96 * se_rl, Upper = rl + 1.96 * se_rl)
}

delta_prof_wave <- function(GEV_samples, n_samples, q_samples, T, fit, residual_sd) {
  
  x_true <- x_samples_sinusoidal(q_samples, fit, residual_sd)
  bm_true <- block_max(x_true, 365)
  true_rl <- as.numeric(quantile(bm_true, probs = 1 - 1/T))

  cat("True return level (T =", T, "):", round(true_rl, 4), "\n")

  # Results storage — one row per simulation
  results <- data.frame(
    delta_rl    = numeric(n_samples),
    delta_lower = numeric(n_samples),
    delta_upper = numeric(n_samples),
    prof_rl     = numeric(n_samples),
    prof_lower  = numeric(n_samples),
    prof_upper  = numeric(n_samples)
  )

  delta_fail     <- 0L
  profile_fail   <- 0L
  profileci_fail <- 0L

  for (i in seq_len(n_samples)) {

    if (i %% max(1, n_samples %/% 4) == 0)
      cat(round(100 * i / n_samples), "% done\n")

    x_dummy  <- x_samples_sinusoidal(GEV_samples, fit, residual_sd)
    bm_dummy <- block_max(x_dummy, 365)

    # Delta method
    fit_delta <- tryCatch(
      suppressWarnings(gev.fit(bm_dummy, show = FALSE)),
      error = function(e) NULL
    )

    if (is.null(fit_delta)) {
      delta_fail <- delta_fail + 1L
      results[i, c("delta_rl", "delta_lower", "delta_upper")] <- NA
    } else {
      ci <- delta_rl_CI(T, fit_delta)
      results[i, "delta_rl"]    <- ci$ReturnLevel
      results[i, "delta_lower"] <- ci$Lower
      results[i, "delta_upper"] <- ci$Upper
    }

    # Profile likelihood
    fit_prof <- tryCatch(
      suppressWarnings(gev_mle(bm_dummy, 1)),
      error = function(e) NULL
    )

    if (is.null(fit_prof) || any(is.na(fit_prof))) {
      profile_fail <- profile_fail + 1L
      results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
    } else {
      prof_return <- gev_return(fit_prof, m = T)
      prof_conf   <- tryCatch(
        suppressWarnings(confint(prof_return, profile = TRUE)),
        error = function(e) NULL
      )

      if (is.null(prof_conf)) {
        profileci_fail <- profileci_fail + 1L
        results[i, c("prof_rl", "prof_lower", "prof_upper")] <- NA
      } else {
        results[i, "prof_rl"]    <- as.numeric(prof_return)
        results[i, "prof_lower"] <- prof_conf[1, 1]
        results[i, "prof_upper"] <- prof_conf[1, 2]
      }
    }
  }

  # Summary
  summary_df <- data.frame(
    ReturnPeriod = T,
    true_rl      = true_rl,

    delta_rl   = mean(results$delta_rl,  na.rm = TRUE),
    prof_rl    = mean(results$prof_rl,   na.rm = TRUE),

    RMSE_delta = sqrt(mean((results$delta_rl - true_rl)^2, na.rm = TRUE)),
    RMSE_prof  = sqrt(mean((results$prof_rl  - true_rl)^2, na.rm = TRUE)),

    Bias_delta     = mean(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof      = mean(results$prof_rl  - true_rl, na.rm = TRUE),

    Bias_delta_med = median(results$delta_rl - true_rl, na.rm = TRUE),
    Bias_prof_med  = median(results$prof_rl  - true_rl, na.rm = TRUE),

    Coverage_delta = mean(results$delta_lower <= true_rl &
                          results$delta_upper >= true_rl, na.rm = TRUE),
    Coverage_prof  = mean(results$prof_lower  <= true_rl &
                          results$prof_upper  >= true_rl, na.rm = TRUE),
    
    delta_width    = mean(results$delta_upper - results$delta_lower, na.rm=TRUE),
    prof_width     = mean(results$prof_upper - results$prof_lower, na.rm=TRUE)
    
  )

  list(
    summary            = summary_df,
    raw                = results,        
    true_rl            = true_rl,
    delta_failures     = delta_fail,
    profile_failures   = profile_fail,
    profileci_failures = profileci_fail,
    delta_failure_rate   = delta_fail     / n_samples,
    profile_failure_rate = profile_fail   / n_samples
  )
}

coverage_plots <- function(raw, true_rl) {
  
  raw     <- raw
  true_rl <- true_rl
  
  n <- nrow(raw)
  raw$iteration <- seq_len(n)
  
  raw$delta_covers <- raw$delta_lower <= true_rl & raw$delta_upper >= true_rl
  raw$prof_covers  <- raw$prof_lower  <= true_rl & raw$prof_upper  >= true_rl
  
  # ── Delta plot ────────────────────────────────────────────────────────────────
  p_delta <- ggplot(raw, aes(x = iteration, colour = delta_covers)) +
  
    geom_linerange(aes(ymin = delta_lower, ymax = delta_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = delta_rl), shape = 16, size = 1.5, alpha = 0.5) +
  
    scale_colour_manual(
      name   = "Covers True Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", hjust = 0.5),
      legend.position  = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin    = margin(4, 6, 4, 6)
    )
  
  # ── Profile plot ──────────────────────────────────────────────────────────────
  p_prof <- ggplot(raw, aes(x = iteration, colour = prof_covers)) +
  
    geom_linerange(aes(ymin = prof_lower, ymax = prof_upper),
                   linewidth = 0.4, alpha = 0.7) +
    geom_point(aes(y = prof_rl), shape = 16, size = 1.5) +
  
    scale_colour_manual(
      name   = "Covers True Return Level",
      values = c("TRUE" = "lightblue", "FALSE" = "blue"),
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    
    geom_hline(yintercept = true_rl,
               linewidth = 0.5, colour = "black") +
    annotate("text", x = 1, y = true_rl,
             label = "True Return Level", hjust = 0, vjust = -0.4, size = 3.2,
             colour = "black") +
    
    scale_x_continuous(
      name   = "Iteration",
      breaks = seq(0, n, by = 100),
      expand = expansion(mult = c(0.02, 0.02))
    ) +
    scale_y_continuous(name = "Return Level") +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor  = element_blank(),
      plot.title        = element_text(face = "bold", hjust = 0.5),
      legend.position   = c(0.18, 0.85),
      legend.background = element_rect(fill = "white", colour = "grey70"),
      legend.margin     = margin(4, 6, 4, 6)
    )
  
  list(
    delta   = p_delta,
    profile = p_prof
  )
}

```

# Location 1 
```{r Location 1 smooth}
load("Red_Sea_data.RData")

x <- data[, 1] # this is only for location 1 

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

residuals <- x - seasonal_mean
resid_sd <- sd(residuals)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365

set.seed(123)
smooth_loc1_redsea_25 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 25, fit_sin, resid_sd)
set.seed(123)
smooth_loc1_redsea_50 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 50, fit_sin, resid_sd)
set.seed(123)
smooth_loc1_redsea_100 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 100, fit_sin, resid_sd)


save(smooth_loc1_redsea_25, 
     smooth_loc1_redsea_50,
     smooth_loc1_redsea_100,
     file = "smooth_loc1_Red_Sea.RData")


```

```{r Location 1 Data}

load("smooth_loc1_Red_Sea.RData")

smooth_loc1_redsea <- rbind(
  smooth_loc1_redsea_25$summary, 
  smooth_loc1_redsea_50$summary,
  smooth_loc1_redsea_100$summary
)

smooth_loc1_redsea

plots <- coverage_plots(
  smooth_loc1_redsea_100$raw, 
  smooth_loc1_redsea_100$summary$true_rl
  )

plots$delta
plots$profile

```


# Location 2
```{r Location 2 smooth}

x <- data[, 2] 

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

residuals <- x - seasonal_mean
resid_sd <- sd(residuals)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365

set.seed(123)
smooth_loc2_redsea_25 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 25, fit_sin, resid_sd)
set.seed(123)
smooth_loc2_redsea_50 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 50, fit_sin, resid_sd)
set.seed(123)
smooth_loc2_redsea_100 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 100, fit_sin, resid_sd)


save(smooth_loc2_redsea_25, 
     smooth_loc2_redsea_50,
     smooth_loc2_redsea_100,
     file = "smooth_loc2_Red_Sea.RData")


```

```{r Location 2 Data}

load("smooth_loc2_Red_Sea.RData")

smooth_loc2_redsea <- rbind(
  smooth_loc2_redsea_25$summary, 
  smooth_loc2_redsea_50$summary,
  smooth_loc2_redsea_100$summary
)

smooth_loc2_redsea

plots <- coverage_plots(
  smooth_loc2_redsea_100$raw, 
  smooth_loc2_redsea_100$summary$true_rl
  )

plots$delta
plots$profile

```

# Location 3
```{r Location 3 smooth}

x <- data[, 3] 

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

residuals <- x - seasonal_mean
resid_sd <- sd(residuals)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365

set.seed(123)
smooth_loc3_redsea_25 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 25, fit_sin, resid_sd)
set.seed(123)
smooth_loc3_redsea_50 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 50, fit_sin, resid_sd)
set.seed(123)
smooth_loc3_redsea_100 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 100, fit_sin, resid_sd)


save(smooth_loc3_redsea_25, 
     smooth_loc3_redsea_50,
     smooth_loc3_redsea_100,
     file = "smooth_loc3_Red_Sea.RData")


```

```{r Location 3 Data}

load("smooth_loc3_Red_Sea.RData")

smooth_loc3_redsea <- rbind(
  smooth_loc3_redsea_25$summary, 
  smooth_loc3_redsea_50$summary,
  smooth_loc3_redsea_100$summary
)

smooth_loc3_redsea

plots <- coverage_plots(
  smooth_loc3_redsea_100$raw, 
  smooth_loc3_redsea_100$summary$true_rl
  )

plots$delta
plots$profile

```


# Location 4
```{r Location 4 smooth}

x <- data[, 4] 

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

residuals <- x - seasonal_mean
resid_sd <- sd(residuals)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365

set.seed(123)
smooth_loc4_redsea_25 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 25, fit_sin, resid_sd)
set.seed(123)
smooth_loc4_redsea_50 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 50, fit_sin, resid_sd)
set.seed(123)
smooth_loc4_redsea_100 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 100, fit_sin, resid_sd)


save(smooth_loc4_redsea_25, 
     smooth_loc4_redsea_50,
     smooth_loc4_redsea_100,
     file = "smooth_loc4_Red_Sea.RData")


```

```{r Location 4 Data}

load("smooth_loc4_Red_Sea.RData")

smooth_loc4_redsea <- rbind(
  smooth_loc4_redsea_25$summary, 
  smooth_loc4_redsea_50$summary,
  smooth_loc4_redsea_100$summary
)

smooth_loc4_redsea

plots <- coverage_plots(
  smooth_loc4_redsea_100$raw, 
  smooth_loc4_redsea_100$summary$true_rl
  )

plots$delta
plots$profile

```

# Location 5
```{r Location 5 smooth}

x <- data[, 5] 

n <- length(x)

day <- 1:n

# Position within the year
doy <- ((day - 1) %% 365) + 1

fit_sin <- lm(
  x ~ sin(2 * pi * doy / 365) +
      cos(2 * pi * doy / 365)
)

seasonal_mean <- predict(fit_sin)

residuals <- x - seasonal_mean
resid_sd <- sd(residuals)

GEV_samples <- 50 * 365
n_samples   <- 1000
q_samples   <- 100000 * 365

set.seed(123)
smooth_loc5_redsea_25 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 25, fit_sin, resid_sd)
set.seed(123)
smooth_loc5_redsea_50 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 50, fit_sin, resid_sd)
set.seed(123)
smooth_loc5_redsea_100 <- delta_prof_wave(GEV_samples, n_samples, q_samples, 100, fit_sin, resid_sd)


save(smooth_loc5_redsea_25, 
     smooth_loc5_redsea_50,
     smooth_loc5_redsea_100,
     file = "smooth_loc5_Red_Sea.RData")


```

```{r Location 5 Data}

load("smooth_loc5_Red_Sea.RData")

smooth_loc5_redsea <- rbind(
  smooth_loc5_redsea_25$summary, 
  smooth_loc5_redsea_50$summary,
  smooth_loc5_redsea_100$summary
)

smooth_loc5_redsea

plots <- coverage_plots(
  smooth_loc5_redsea_100$raw, 
  smooth_loc5_redsea_100$summary$true_rl
  )

plots$delta
plots$profile

```


# THE FAT TABLE FOR LOCATIONS 1 to 5 
```{r Table code}

load("smooth_loc1_Red_Sea.RData")

smooth_loc1_redsea <- rbind(
  smooth_loc1_redsea_25$summary, 
  smooth_loc1_redsea_50$summary,
  smooth_loc1_redsea_100$summary
)

load("smooth_loc2_Red_Sea.RData")

smooth_loc2_redsea <- rbind(
  smooth_loc2_redsea_25$summary, 
  smooth_loc2_redsea_50$summary,
  smooth_loc2_redsea_100$summary
)

load("smooth_loc3_Red_Sea.RData")

smooth_loc3_redsea <- rbind(
  smooth_loc3_redsea_25$summary, 
  smooth_loc3_redsea_50$summary,
  smooth_loc3_redsea_100$summary
)

load("smooth_loc4_Red_Sea.RData")

smooth_loc4_redsea <- rbind(
  smooth_loc4_redsea_25$summary, 
  smooth_loc4_redsea_50$summary,
  smooth_loc4_redsea_100$summary
)


load("smooth_loc5_Red_Sea.RData")

smooth_loc5_redsea <- rbind(
  smooth_loc5_redsea_25$summary, 
  smooth_loc5_redsea_50$summary,
  smooth_loc5_redsea_100$summary
)


prepare_table_prof <- function(x, scenario, method){
  
  if(method == "Profile"){
    
    x |>
      select(
        ReturnPeriod,
        true_rl,
        prof_rl,
        RMSE_prof,
        Bias_prof,
        Bias_prof_med,
        Coverage_prof,
        prof_width
      ) |>
      rename(
        T = ReturnPeriod,
        `True RL` = true_rl,
        `Est RL` = prof_rl,
        RMSE = RMSE_prof,
        Bias = Bias_prof,
        `MD Bias` = Bias_prof_med,
        Coverage = Coverage_prof,
        Width = prof_width
      )
  } |>
    mutate(
      Scenario = scenario,
      .before = 1
    )
}


prof_all <- rbind(
  prepare_table_prof(smooth_loc1_redsea, "RS1", "Profile"),
  prepare_table_prof(smooth_loc2_redsea, "RS2", "Profile"),
  prepare_table_prof(smooth_loc3_redsea, "RS3", "Profile"),
  prepare_table_prof(smooth_loc4_redsea, "RS4", "Profile"),
  prepare_table_prof(smooth_loc5_redsea, "RS5", "Profile")
)


prof_all <- prof_all |>
  mutate(
    across(
      where(is.numeric),
      ~ round(.x, 3)
    )
  )

prof_all |>
  kbl(
    format = "latex",
    booktabs = TRUE,
    caption = "Profile likelihood return level estimation performance for the smooth harmonic Red Sea simulation Locations 1 to 5. True return levels are obtained from the 100,000 year reference sample generated under the same weekly mean/sd structure.",
    label = "tab:estimation"
  ) |>
  save_kable("smooth_loc12345_RS_table.tex")



```
